---
title: "Data Privacy Survey: signalling challenges and improvements of privacy-related services for research"
author: "Dorien Huijser, Neha Moopen"
date: "26 april 2022"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

## Abstract
In the second quarter of 2022, Utrecht University (UU) Research Data Management 
Support sent out a survey among all scientific personnel at UU: the Data Privacy
Survey. The aim of this survey was to investigate 1) How UU researchers 
currently deal with personal data in research, 2) What challenges they run into 
when handling personal data in research, and 3) How RDM Support and 
collaborators can improve their services and support concerning personal data 
in research. The current report contains a summary of the results of this 
survey, as well as points of action to be taken.

## Report contents

1. Management summary
2. Introduction
3. Methods
4. Results
5. Conclusion and discussion
6. References
7. Technical information

## Management summary

## Introduction
Scientific research often includes some form of personal data. The amount of 
personal data in research is likely underestimated, as researchers may be 
unaware of what personal data are or whether they are being collected. With the 
implementation of the General Data Protection Regulation (GDPR) in May 2018, 
stricter legal requirements apply to handling personal data and its sharing or 
publication. The number of (complex) questions on handling personal data in 
scientific research is increasing. Yet, practical information to resolve these 
questions can be difficult to find, scattered, and/or discipline-specific. As 
such, there is room to better translate the legislation to scientific research 
practice and to provide practical tools for handling personal data in research.

This room for better translation is exemplified in the 
<a href = "https://utrechtuniversity.github.io/dataprivacyproject/" 
target = "_blank">Data Privacy Project</a>, a data support effort^[The Data 
Privacy Project is led by Utrecht University's <a href="https://uu.nl/rdm/" 
target = "_blank">Research Data Management Support</a>,
a collaboration between the University Library and Information Technology 
Services, but there is a tight collaboration with other data support, such as 
first- and second-line privacy officers and information security.] within 
Utrecht University (UU), the Netherlands. The project aims to build knowledge 
and expertise on how researchers can and should deal with personal data. The 
project's deliverables are all focussed on providing practical, actionable and
FAIR (Findable, Accessible, Interoperable, Reusable) information and tools for
researchers to use in their research ^[The project is funded by Utrecht 
University’s Research IT program and a Digital Competence Center grant from 
the Dutch Organization for Scientific Research (NWO).]. 

To properly take into account the challenges and needs of our target audience, 
namely UU researchers, one of the deliverables of the project is to perform
quantitative and qualitative investigations into the current practices and needs
of UU researchers concerning the handling of personal data. Our primary research
questions are:

1. To what extent are UU researchers aware of privacy legislation and practices?
2. What data privacy issues do UU researchers typically run into?
3. What support do UU researchers need to handle personal data?

To answer these questions both quantitatively and qualitatively, we have set 
up an online survey and planned one-on-one meetings with a selection of UU 
researchers. This report describes our methods, results and the future steps
that data support at UU can take to improve their services concerning privacy
in research.

## Methods
### Data Privacy Survey
```{r dependencies, warning = FALSE, message = FALSE}
# install.packages("data.table")
library(data.table)

# install.packages("tidyverse")
library(tidyverse)

#install.packages("wordcloud2")
library(wordcloud2)

# install.packages("gridExtra")
library(gridExtra)
```

```{r readdata, cache=TRUE}
# This file has multiple responses in separate columns
datafile <- "../data/raw/Data_Privacy_Survey_20220502.csv"
dppsurvey <- fread(datafile, na.strings = "")

# Skip the first 2 rows as they contain the question text
dppsurvey <- dppsurvey[-c(1, 2), ]

# Skip the previews
dppsurvey <- dppsurvey[-grep("preview",
                             dppsurvey$DistributionChannel), ]

# Convert Duration and Progress into a numeric variable
dppsurvey$Duration <- as.numeric(dppsurvey$`Duration (in seconds)`)
dppsurvey$Progress <- as.numeric(dppsurvey$Progress)

# Skip unnecessary columns
dppsurvey <- dppsurvey[,-c("StartDate","EndDate","Status","IPAddress", 
                           "Duration (in seconds)", "Finished", "RecordedDate", 
                           "ResponseId", "RecipientLastName", "RecipientFirstName",
                           "RecipientEmail", "ExternalReference", "LocationLatitude",
                           "LocationLongitude", "DistributionChannel", "UserLanguage")]

# Select only consented data
dppsurvey <- dppsurvey %>% 
  filter(Consent == "Yes")
```

#### Survey questions 
The Data Privacy Survey was created by the project coordinators (DH, NM) with 
input from the Data Privacy Project team, consisting of privacy officers, data
managers, data consultants, information security and research engineering.
The final version of the survey in Qualtrics consisted of 19-23 questions, 
which were separated into 5 sections:

1. *About you and your research*: 5 questions about the faculty, department, 
position, and type of (personal) data the researcher works with.
2. *Measures and documents*: 4-6 questions about researchers' familiarity and
use of privacy-related measures (e.g., processing register, Data Privacy Impact
Assessment, encryption, pseudonymisation, etc.), data storage, and informed 
consent forms.
3. *Data sharing, archiving and publication*: 3-4 questions about researchers' 
data sharing and publishing practices.
4. *Finding support*: 3-4 questions about the awareness of several data support
channels.
5. *Improving our services*: 4 questions about researchers' challenges, needs, 
and suggestions to improve UU's data privacy support.

The estimated 10 minute duration of the survey in advance proved to be 
relatively accurate: respondents spent an average of 
`r round(median(dppsurvey$Duration/60), 1)` minutes on the survey (range: 
`r round(min(dppsurvey$Duration)/60, 1)` minutes - 
`r round(max(dppsurvey$Duration)/3600/24,1)` days). The majority of researchers
(`r round(length(dppsurvey$Progress[dppsurvey$Progress == 100])/dim(dppsurvey)[1]*100,1)`%) 
fully completed the survey.

#### Procedure
The survey was distributed from March 21st, 2022 onwards via several 
communication channels to reach as many UU researchers as possible:

- An <a href = "https://utrechtuniversity.github.io/dataprivacyproject/20220321-data-privacy-survey.html"
target = "_blank">email</a> was sent to all academic staff at UU through 
central communication channels. The reasoning was that this would be the most
effective way to reach as many UU researchers as possible (taking for granted
that we would likely miss a small part of non-academic personnel also involved
in research in some way).
- A mention in several Faculty newsletters.
- Social media messages (e.g., on <a href = "https://twitter.com/RDMsupportUU/status/1517390979707412481"
target = "_blank">Twitter</a>).
- A <a href = "https://intranet.uu.nl/en/news/news-items/survey-better-support-for-handling-privacy-sensitive-data-in-research" 
target = "_blank">news item</a> on the UU intranet.
- Via data support colleagues, who were asked to point researchers they were
in contact with to the survey.

Researchers could provide voluntarily whether they wanted to participate in the
survey. The survey only started after they had provided their consent. All 
results reported here are from researchers who provided their active consent. 

#### Survey respondents
The survey was filled out by `r dim(dppsurvey)[1]` UU researchers. The survey
was created in such a way so that only researchers from a UU faculty, who 
worked with personal research data, were selected to respond to all other survey
questions.

We received responses from each faculty, but the distribution is not equal: 
the faculties of Science and Social and Behavioural Sciences are 
overrepresented, whereas the responses of the Faculty of Geosciences and the 
Faculty of Medicine are rather low. This can be explained by the fact that 
a large chunk of research performed at the  Faculty of Geosciences does 
not investigate human behavior and therefore does not work with personal data 
much. The Faculty of Medicine, then, is located at the University Medical 
Center Utrecht (UMCU) and is not of primary interest for the current survey, 
as it concerns an entirely different organisation with a separate research 
support structure.

As can be seen in the figure on the right, most respondents are early 
career researchers (PhDs, junior researchers and postdoctoral researchers). 
We suspect that this is because 1) there simply are more early-career 
researchers than mid- and late career researchers, and 2) early-career 
researchers experience a greater need for support with respect to handling 
personal data.

```{r facultyplot}
# R color brewers palettes: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
# UU colors: https://www.uu.nl/organisatie/huisstijl/huisstijlelementen/kleur
UU_pallette <- c("#FFCD00", "#FFE6AB", "#F3965E", "#AA1555", "#6E3B23", "#24A793", "#5287C6", "#001240", "#5B2182")

# Simple barplot of faculty division (WORKS, yet to prettify)
facultyplot <- 
  dppsurvey %>% 
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty) %>%
  summarise(Count = length(Faculty)) %>%
  arrange(Count) %>% # Order by count
  mutate(Faculty = factor(Faculty, levels=Faculty)) %>% # Update factor levels
  ggplot(aes(x = Faculty, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Faculties",
       caption = "UU faculty representation in the Data Privacy Survey") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Simple barplot of position division (WORKS, yet to prettify)
positionplot <-
  dppsurvey %>% 
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Position) %>%
  summarise(Count = length(Position)) %>%
  arrange(Count) %>% # Order by count
  mutate(Position = factor(Position, levels=Position)) %>% # Update factor levels
  ggplot(aes(x = Position, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Position",
       caption = "Positions of Data Privacy Survey respondents") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(facultyplot, positionplot, nrow = 1)
```

#### Analysis
Below we report on the descriptive statistics or summaries from most of the 
survey questions. As we did not formulate hypotheses, no statistical analyses 
were performed. 

#### Data and material availability
The entire survey can be found 
<a href = "https://utrechtuniversity.github.io/dataprivacyproject/assets/docs/survey-questions-qualtrics.pdf"
target = "_blank">on our website</a>, as well as the separate 
<a href = "https://utrechtuniversity.github.io/dataprivacyproject/assets/docs/survey-privacy-statement.pdf" 
target = "_blank">privacy statement</a>. The script can be found in the designated
<a href = "https://github.com/UtrechtUniversity/dataprivacysurvey" target = 
"_blank">GitHub repository</a>. As the raw data contain personal information (IP 
address, demographic information, open text responses and email addresses), and
no consent was obtained to share those details, we are unable to share them.

### One-on-one meetings
The one-on-one meetings are currently being conducted. More will be shared soon!

## Results

### Types of research data

Within the sample of the survey, it is clear that most researchers use tabular, textual, code and audio data as their primary research data format. The most encountered type of *personal* data that researchers indicate to work with are contact information, demographic information and direct identifiers. This can be either because these are the types of personal data that are indeed most common, or because researchers mostly recognize these types of data as being personal data.

```{r datatypes}
# count
#sum(!is.na(dppsurvey$Datatype_1))

# percentage
#sum(!is.na(dppsurvey$Datatype_1))/dim(dppsurvey)[1]

# text label
#unique(dppsurvey$Datatype_1)[1]

# Table of data types
datatypetable <- dppsurvey %>% 
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

# Table of personal data types
tablepersdata <- dppsurvey %>% 
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

# Donut chart of personal data types --> Nr of categories = too large! But does work otherwise
#perstypedonut <- dppsurvey %>% 
#  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
#               values_drop_na = TRUE,
#               values_to = "Personal_Datatype") %>%
#  select(-name) %>%
#  group_by(Personal_Datatype) %>%
#  summarise(Count = length(Personal_Datatype)) %>%
#  arrange(Count) %>% # Order by count
#  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
#         Percentage = round(Count/dim(dppsurvey)[1]*100,1),
#         ymax = cumsum(Percentage),
#         ymin = c(0, head(ymax, n = -1)),
#         labelPosition = (ymax + ymin)/2,
#         label = paste0(Personal_Datatype, ": ", Percentage, "%")) %>% 
#  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = Personal_Datatype)) +
#  geom_rect() + 
#  geom_label(x = 3.5, aes(y = labelPosition, label = label), size = 6) +
#  scale_fill_brewer(palette=4) +
# coord_polar(theta="y") + 
# xlim(c(2,4)) + 
#  theme_void() +
#  theme(legend.position = "none")
```

````{r todo}
# **TO DO: wrangle Dept_ columns so that they can be displayed within Faculties**
````

### Current practices in handling personal data

The first part of the survey addressed the researchers' current practices in 
handling personal data in their research.

#### - Organisational and technical measures

With respect to organisational and technical measures used to handle personal data, most researchers indicate they pseudonymise/anonymise their data. We cannot, however, assess whether the data is sufficiently pseudonymised/anonymised in line with the GDPR. Researchers also implement some form of access control and complete data management plans (DMPs). Agreements, such as NDAs, are utilized to some degree as well. On the other hand, GDPR-specific assessments such as DPIAs or privacy reviews are least used. As it stands, these assessments are typically carried out on a case-by-case basis.   
 
```{r measuresused}

# table of measures used

hpd_measures_used <- datatypetable <- dppsurvey %>% 
  pivot_longer(cols = grep("^Orgmeasures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "measures") %>%
  select(-name) %>%
  group_by(measures) %>%
  summarise(Count = length(measures)) %>%
  arrange(Count) %>% # Order by count
  mutate(measures = factor(measures, levels=measures)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()
```

#### Storage media used

To store personal data, most researchers rely on storage solutions provided by UU. Nonetheless, some researchers utilize non-UU solutions, including non-UU cloud storage. Several researchers also indicated they use 'Other' storage solutions - the survey responses do not provide more information on this option.

```{r storagemedia}
# table storage media used

storage_used <- dppsurvey %>% 
  pivot_longer(cols = grep("^Storage_medium_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "storage") %>%
  select(-name) %>%
  group_by(storage) %>%
  summarise(Count = length(storage)) %>%
  arrange(Count) %>% # Order by count
  mutate(storage = factor(storage, levels=storage)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()
```

### Consent forms: usage and content

Only a small number of researchers indicated that consent forms were not applicable to their research or they didn't know what they were. When researchers do use consent forms, they often use forms approved by privacy officers or ethical committees. Researchers also indicated they tend to independently create and use forms. Departmental and UU-wide templates are also used to a fair extent, as well as 'Other' templates.



```{r consentforms}
# table of consent forms used

consent_forms <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consentforms_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "consent_forms") %>%
  select(-name) %>%
  group_by(consent_forms) %>%
  summarise(Count = length(consent_forms)) %>%
  arrange(Count) %>% # Order by count
  mutate(consent_forms = factor(consent_forms, levels=consent_forms)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()
```

```{r consentcontent}
# table of content in consent forms

consent_content <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consent_content_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "consent_content") %>%
  select(-name) %>%
  group_by(consent_content) %>%
  summarise(Count = length(consent_content)) %>%
  arrange(Count) %>% # Order by count
  mutate(consent_content = factor(consent_content, levels=consent_content)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()
```

### DPIA: familiarity and help used

```{r dpia}
# table  on experience with DPIAs

DPIA_experience <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_experience_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "DPIA_experience") %>%
  select(-name) %>%
  group_by(DPIA_experience) %>%
  summarise(Count = length(DPIA_experience)) %>%
  arrange(Count) %>% # Order by count
  mutate(DPIA_experience = factor(DPIA_experience, levels=DPIA_experience)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

# table on help with DPIAs

DPIA_help <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_Help_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "DPIA_help") %>%
  select(-name) %>%
  group_by(DPIA_help) %>%
  summarise(Count = length(DPIA_help)) %>%
  arrange(Count) %>% # Order by count
  mutate(DPIA_help = factor(DPIA_help, levels=DPIA_help)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

### Data sharing practices: externally and if so, which measures are used?

```{r datasharing}
external_sharing <- dppsurvey %>% 
  pivot_longer(cols = grep("^Share_outside_UU_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "external_sharing") %>%
  select(-name) %>%
  group_by(external_sharing) %>%
  summarise(Count = length(external_sharing)) %>%
  arrange(Count) %>% # Order by count
  mutate(external_sharing = factor(external_sharing, levels=external_sharing)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

```{r sharingmeasures}
sharing_measures <- dppsurvey %>% 
  pivot_longer(cols = grep("^Share_measures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "sharing_measures") %>%
  select(-name) %>%
  group_by(sharing_measures) %>%
  summarise(Count = length(sharing_measures)) %>%
  arrange(Count) %>% # Order by count
  mutate(sharing_measures = factor(sharing_measures, levels=sharing_measures)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

### Data publishing

Does that ever happen and if so, in what format. If not: why not?

```{r datapublishing}
data_publishing <- dppsurvey %>% 
  pivot_longer(cols = grep("^Data_publication_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "data_publishing") %>%
  select(-name) %>%
  group_by(data_publishing) %>%
  summarise(Count = length(data_publishing)) %>%
  arrange(Count) %>% # Order by count
  mutate(data_publishing = factor(data_publishing, levels=data_publishing)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

```{r pubformat}
# publication format

publication_format <- dppsurvey %>% 
  pivot_longer(cols = grep("^Publication_format_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "publication_format") %>%
  select(-name) %>%
  group_by(publication_format) %>%
  summarise(Count = length(publication_format)) %>%
  arrange(Count) %>% # Order by count
  mutate(publication_format = factor(publication_format, levels=publication_format)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

```{r reasonsnopub}
# reasons for not publishing

reasons_nopub <- dppsurvey %>% 
  pivot_longer(cols = grep("^Reason_NoPub_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "reasons_nopub") %>%
  select(-name) %>%
  group_by(reasons_nopub) %>%
  summarise(Count = length(reasons_nopub)) %>%
  arrange(Count) %>% # Order by count
  mutate(reasons_nopub = factor(reasons_nopub, levels=reasons_nopub)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()

```

### Existing support channels

#### Faculty privacy officer
The second part of the survey concerned the visibility and use of existing 
support channels. When asked whether respondents know who their faculty
privacy officer is, most respondents indicated that they do 
not (Yes: `r sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/dim(dppsurvey)[1]*100,0)`%), No: 
`r sum(dppsurvey$Know_PO=="No", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/dim(dppsurvey)[1]*100,0)`%), NA: 
`r sum(is.na(dppsurvey$Know_PO))` 
(`r round(sum(is.na(dppsurvey$Know_PO))/dim(dppsurvey)[1]*100,0)`%)). This 
suggests either that researchers have simply never required help from their privacy 
officer, or that the faculty privacy officers could increase their visibility
within their faculties.

#### Looking for help

When asked whether respondents have ever looked for support in the form of 
information, tools, or support an overwheling majority indicated that they have,
as can be seen in the graph below. Most respondents that looked for support
indicate however that they haven't always found the suport they were looking for.
Together with the results from the previous question about privacy officers,
this suggests that the visibility of the current support channels could be 
improved.

```{r existingsupport}
# Ever looked for help with privacy?
lookedforhelpplot <- dppsurvey %>% 
  pivot_longer(cols = Lookedforhelp, 
               values_drop_na = TRUE,
               values_to = "searchhelp") %>%
  select(-name) %>%
  group_by(searchhelp) %>%
  summarise(Count = length(searchhelp)) %>%
  arrange(Count) %>% # Order by count
  mutate(searchhelp = factor(searchhelp, levels = searchhelp)) %>% # Update factor levels
  ggplot(aes(x = searchhelp, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Has looked for help",
       caption = "Have you ever looked for UU-specific information, \n in-person support or tools on handling personal data?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Found what you're looking for?
foundhelpplot <- dppsurvey %>% 
  pivot_longer(cols = Foundhelp, 
               values_drop_na = TRUE,
               values_to = "foundhelp") %>%
  select(-name) %>%
  group_by(foundhelp) %>%
  summarise(Count = length(foundhelp)) %>%
  arrange(Count) %>% # Order by count
  mutate(foundhelp = factor(foundhelp, levels = c("(Almost) always", 
                                                  "Often",
                                                  "Sometimes",
                                                  "(Almost) never"))) %>% # Update factor levels
  ggplot(aes(x = foundhelp, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Has found what they were looking for",
       caption = "Did you find what you were looking for \n(e.g., information, support, tools)?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(lookedforhelpplot, foundhelpplot, nrow = 1)
```

#### Channels used to find support

The graph below indicates which channels respondents to the survey used most 
when looking for information about handling personal data. Note that this was 
a multiple choice question, and so the numbers do no add up to the total number
of survey respondents. As shown below, all sources mentioned are to some extent
consulted by the respondents. However, there is large differentiation between 
these sources. The university website and intranet are the most visited online 
resources for information about handline personal data. Notably, colleagues 
appear to play an important role as well in informing researchers
about how to handle personal data.

```{r sourcesused}
# Sources used to find information and help
sourcesusedplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Sourcesused_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "sourcesused") %>%
  select(-name) %>%
  group_by(sourcesused) %>%
  summarise(Count = length(sourcesused)) %>%
  arrange(Count) %>% # Order by count
  mutate(sourcesused = factor(sourcesused, levels = sourcesused)) %>% # Update factor levels
  ggplot(aes(x = sourcesused, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Sources used to find help") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

sourcesusedplot

# Other sources used
sourcesusedother <- 
  dppsurvey$Sourcesused_19_TEXT[!is.na(dppsurvey$Sourcesused_19_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()
```


### Challenges and needs
```{r obstacle}
obstacleplot <- dppsurvey %>% 
  pivot_longer(cols = Obstacle, 
               values_drop_na = TRUE,
               values_to = "obstacle") %>%
  select(-name) %>%
  group_by(obstacle) %>%
  summarise(Count = length(obstacle)) %>%
  arrange(Count) %>% # Order by count
  mutate(obstacle = factor(obstacle, levels = c("(Almost) always",
                                                "Sometimes",
                                                "(Almost) never",
                                                "Not sure"))) %>% # Update factor levels
  ggplot(aes(x = obstacle, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Is privacy an obstacle?",
       caption = "Do you see dealing with personal data as an obstacle in open science and research data management?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

obstacleplot
```

As can be seen above, most researchers experience privacy an obstacle for 
open science and research data management. It is therefore important to aim for
better service and support in this area. What this support should look like 
differs a bit: 

```{r bettersupport}
bettersupporttable <- dppsurvey %>% 
  pivot_longer(cols = grep("^Better_support_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "bettersupport") %>%
  select(-name) %>%
  group_by(bettersupport) %>%
  summarise(Count = length(bettersupport)) %>%
  arrange(Count) %>% # Order by count
  mutate(bettersupport = factor(bettersupport, levels=bettersupport)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  print()
```

```{r selectopenquestions}
opentextresponses <- dppsurvey %>% 
  select(
    contains(c("Faculty", "Dept", "Position")),
    Better_support_9_TEXT,
    Challenges,
    Infotools_missing
  ) %>%
  unite("Faculty", grep("Faculty",names(.), value = T), 
        sep = ", ", remove = TRUE, na.rm=TRUE) %>%
  unite("Dept", grep("Dept",names(.), value = T), 
        sep = ", ", remove = TRUE, na.rm=TRUE) %>%
  unite("Position", grep("Position",names(.), value = T), 
        sep = ", ", remove = TRUE, na.rm=TRUE) %>%
  rename(Bettersupport = Better_support_9_TEXT) %>%
  mutate(
    Bettersupport = str_to_lower(str_trim(Bettersupport)),
    Challenges = str_to_lower((str_trim(Challenges))),
    Infotools_missing = str_to_lower(str_trim(Infotools_missing))
  ) %>%
  filter_at(vars(Bettersupport, 
                 Challenges, 
                 Infotools_missing), 
            any_vars(!is.na(.)))

todaysdate <- as.character(Sys.Date())
write.csv(opentextresponses, file.path(paste0("../data/raw/", todaysdate, "_dppsurvey_opentextresponses.csv")))
```


The open text responses reveal the following challenges and needs. Please note that some researchers indicate not to have run into any issues, but for action point determination purposes, these have been left out here:

- **Time-intensive and bureaucratic**: Processes are often experienced as being time-inefficient, and sometimes longer and more bureaucratic than necessary. Some researchers argue that (part of) this burden should be relieved or carried by support staff:
<blockquote>"Sharing data costs a lot of time and is inefficient when you do not do it often"</blockquote>
<blockquote>"Minder acties die gericht zijn op inhoudelijk trainen van wp en meer uit handen nemen van deze groep."</blockquote>
- **Unclear expectations and support channels**: There is unclarity in the exact requirements for researchers: what does the process look like? Who should they ask for help, and with what kinds of questions?
<blockquote>"There is no one who tells at the start of your PhD how you should handle your data. [...] I think new PhD students should get a basic course on data management and privacy."</blockquote>
- **More hands-on support**: Some researchers find that support staff could provide support in a more hands-on fashion, rather than abstract advice and telling researchers how *not* to do things. Some researchers add that privacy professionals have a tendency to cling to the letter of the law, instead of looking at how to apply the law in practice:
<blockquote>"I find that there is a severe lack of trust in employees"</blockquote>
<blockquote>"Soms zijn we door deze regels heiliger dan de paus."</blockquote>
- **Specific data types**: Some researchers express the need for practical guidelines, tools and templates on specific types of data (e.g., etnographic, historic, linguistic, video data, etc.) or specific scenarios. Multiple researchers indicate that existing solutions are not always suitable for their type of research, or that there are no (standardised) guidelines available at all.
<blockquote>"De informatie is gewoon veel te generiek, er zouden templates moeten zijn per type onderzoek."</blockquote>
<blockquote>"Veel templates en uitlegmodellen spreken over data, data packages en metadata, maar die woorden zijn niet ingebed in historisch onderzoek. Er ontstaat al snel verwarring over wat historici nu precies moeten met archiefmateriaal in het licht van privacy."</blockquote>
- **Definitions**: There is unclarity among some researchers on when data are personal and when  they are anonymous, and sometimes researchers acknowledge that it is difficult or impossible to anonymise data and then wonder what to do in that case. 
<blockquote>"Ik heb niet zo'n goed beeld van wat precies 'gevoelige' informatie is en wat niet."</blockquote>
<blockquote>"How far should I go in anonymising data?"</blockquote>
- **Specific tools**: Some researchers express a need for a concrete (improvement of a) tool, such as for pseudonymising video data, transcribing audio data, and sharing data with others. 


## Results summary [preliminary]

In this survey, we find that personal data are processed in each faculty and 
across all academic positions at Utrecht University. Most researchers have 
tabular, textual, code and audiovisual datasets, which contain all kinds of 
personal data (mostly demographics, contact information, direct identifiers).
  
Most researchers are familiar with concepts like anonymisation and 
pseudonymisation, access control, and UU-approved tools for handling personal 
data. However, the knowledge level differs quite a bit. For example, some 
researchers seem to store data on locations that officially are not allowed.
Data is often shared, but not all the right measures are always taken to do so 
securely. Additionally, some measures to securely handle personal data in accordance 
with the GDPR are less known, or there is confusion about it. For example, 
the processing register, Data Protection Impact Assessment, and Standard 
Contractual Clauses are among the least known concepts that we measured. And there
seems to be unclarity among part of the respondents on when sharing data is allowed
or when data are personal. Together, this suggests that there is a lot to be gained
regarding awareness and secure practices in scientific research with personal data.

Below, we summarise challenges and needs explicated by researchers. 

## Recommendations [preliminary]
### Increased visibility

Our data show that the visibility of existing support for handling personal data 
can be increased at UU. This includes visibility of privacy officers, legal affairs 
and Research Data Management Support, as well as faculty- or department-level 
support. 

As shown by the fact that many respondents consult with colleagues for 
information and support, personal contact with supporters will likely be more 
effective at increasing awareness and uptake of secure privacy practices than a 
passive website that may not always answer all possible questions.

Some suggestions for increasing visibility:

- Make sure there is (sufficient) support available to help researchers in time.
If there is no support available, the odds are that researchers may not ask for 
help at all and thus forget important privacy elements in the design of their 
research project.
- Make sure support staff are easy to find. For example, give them a clear location
on the university website and/or on the intranet. Also indicate who is available
for what type of support and for which target audience (e.g., "all UU 
researchers", "employees from the Faculty of Humanities", etc.)
- Make sure information on handling personal data in research is easy to find.
The UU website and intranet seem the most suitable places, as these are the most
visited to look for information according to the survey. 
- Make researchers actively aware of existing support channels. For example,
first-line support can visit important faculty or department events, offer to
give a workshop, and keep short lines of contact with department heads.

### More hands-on support

The survey and interviews show that some researchers are not content with the 
support they have received in the past. For example, their research project may
have suffered significant delays due to the fact that support personnel only
pointed them at what they could *not* do, rather than how to solve privacy-
related issues. We noticed many researchers that argue for more hands-on suppport,
rather than advise with no concrete direction how to apply it in practice. More
hands-on support can for example be reached by:

- More collaboration between different types of data support. For example, at UU
there is a clear distinction between privacy officers (employed by faculties),
data managers (employed by the University Library and some faculties), information
security officers (Information Technology Services) and research engineers 
(Information Technology Services). However, in practice there is often overlap 
between either or all of these positions. Regular collaboration between these 
parties can:
  - help privacy and security experts understand the perspective of the 
  researcher and possibly how to apply their advice in practice,
  - help data managers and research engineers understand the GDPR requirements 
  and translate those to practice, and
  - enable reuse of solutions already applied in earlier cases that either 
  party may not be aware of.
- Together with researchers, developing suitable practical solutions (tools, 
protocols, etc.) for specific scenarios and/or types of research, and sharing 
them with the world for reuse (if possible).
- Increasing contact between support staff and researchers to understand the 
issues they are dealing with that require a solution.

### Decreased administrative burden

A selection of researchers mention a high administrative burden when handling
personal data. This is for example expressed in processes that take too long
(e.g., writing and reviewing a Data Privacy Impact Assessment) and too many 
forms to be filled out with overlapping content (e.g., privacy review, Data 
Management Plan, processing register, DPIA). We suggest that these issues can 
be partly attributed to:

1. Understaffing: it is currently impossible for data support staff to review 
all research projects involving personal data that are undertaken at UU. This 
can lead to projects with issues being noticed too late and researchers having 
to change their timeline or even project as a whole after the fact. This creates
unnecessary administrative burden that could have been prevented had researchers
been in contact with support staff. 
2. Lack of unifying policy on steps to take when working with personal data. At
the moment each faculty has its own procedure, or has none at all. This can 
create confusion among both researchers and support staff on which steps to take
when, who to contact for help, or who is responsible for deciding on compliance.
3. Lack of integration between systems and staff: having to fill out the same
thing multiple times may be caused by the fact that different administrative
systems are used for different purposes. Integration of these systems may lead
to a relief of the administrative burden experienced by researchers. <br>For 
example, DMPonline is used at UU for registering Data Management Plans. However, 
DMP templates do not include sufficient questions about how personal data are 
handled, and thus a privacy plan or review may be required. Privacy is 
sometimes part of an ethical application, and so this same information may be 
included there again as well. Finally, the processing register also requires 
information on (among others) the type and sensitivity of personal data, 
which is a third system where this information will be registered.

### Education and outreach

Many problems seem to arise from the fact that privacy has not been taken into
account from the design phase of a research project. This leads to research
support staff being contacted too late in the process, who then has to tell the
researcher to retrospectively make changes in their design in order to move 
forward. 

Such scenarios may be prevented by, for example:

- Including the basic of handling personal data and contact details of support
staff in onboarding processes when researchers are newly hired.
- Educating researchers on taking privacy into account in the designing phase
of their project. 
- Organising outreach events, such as the "Privacy and Security week" making
researchers aware of privacy-related measures (see further Visibility).
- Embedding privacy in existing materials and processes, such as in ethical
applications, in the education of research master and PhD students, etc.

## Conclusion [preliminary]


## References

## Technical information
This document was last created on: `r as.character(Sys.Date())`.

```{r sessioninfo}
sessionInfo()
```
