---
title: "survey-first-glance"
author: "Dorien Huijser"
date: "8 April 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**In the second quarter of 2022, Utrecht University (UU) Research Data Management Support sent out a survey among all scientific personnel at UU: the Data Privacy Survey. The aim of this survey was to investigate 1) How UU researchers currently deal with personal data in research, 2) What challenges they run into when handling personal data in research, and 3) How RDM Support and collaborators can improve their services and support concerning personal data in research. The current report contains a summary of the results of this survey, as well as points of action to be taken.**

## Analysis plan

1. Overall (no split per faculty or position, entire UU): counts, percentages, graphs
2. Split by faculty, possibly department (for organization purposes): counts, percentages, graphs
3. Generate a first report of the report to share and present. Ask for input: what else do people want to know? Ideas for other ways to analyse?
4. After feedback, possible split by Position (for teaching purposes).
5. If relevant, split per faculty and position (if not too small of a sample)

### Useful links for prettifying

- https://www.data-to-viz.com/ 
- https://r-graph-gallery.com/ 
- https://holtzy.github.io/Pimp-my-rmd/ 
- Coolers: color palette with UU colors

```{r dependencies}
# install.packages("data.table")
library(data.table)

# install.packages("tidyverse")
library(tidyverse)

#install.packages("wordcloud2")
library(wordcloud2)

# install.packages("gridExtra")
library(gridExtra)
```

```{r readdata, cache=TRUE}
# We first read the data from Qualtrics and select the relevant data
# This file has multiple responses in separate columns
datafile <- "../data/raw/Data_Privacy_Survey_20220411.csv"
dppsurvey <- fread(datafile, na.strings = "")

# Skip the first 2 rows as they contain the question text
dppsurvey <- dppsurvey[-c(1, 2), ]

# Skip the previews
dppsurvey <- dppsurvey[-grep("preview",
                             dppsurvey$DistributionChannel), ]

# Convert Duration and Progress into a numeric variable
dppsurvey$Duration <- as.numeric(dppsurvey$`Duration (in seconds)`)
dppsurvey$Progress <- as.numeric(dppsurvey$Progress)

# Skip unnecessary columns
dppsurvey <- dppsurvey[,-c("StartDate","EndDate","Status","IPAddress", 
                           "Duration (in seconds)", "Finished", "RecordedDate", 
                           "ResponseId", "RecipientLastName", "RecipientFirstName",
                           "RecipientEmail", "ExternalReference", "LocationLatitude",
                           "LocationLongitude", "DistributionChannel", "UserLanguage")]

# Select only consented data
dppsurvey <- dppsurvey %>% 
  filter(Consent == "Yes")

```


## Data exploration

The survey was filled out by `r dim(dppsurvey)[1]` researchers. The median time it took to complete the survey was `r round(median(dppsurvey$Duration/60), 1)` minutes (range: `r round(min(dppsurvey$Duration)/60, 1)` minutes -`r round(max(dppsurvey$Duration)/3600/24,1)` days). The majority of researchers (`r round(length(dppsurvey$Progress[dppsurvey$Progress == 100])/dim(dppsurvey)[1]*100,1)`%) fully completed the survey.

We received responses from each faculty, but the distribution is not equal: the faculties of Science and Social and Behavioural Sciences are overrepresented, whereas the responses of the Faculty of Geosciences and the Faculty of Medicine are rather low. This can be explained by the fact that a large chunk of research performed at the  Faculty of Geosciences does not investigate human behavior and therefore does not work with personal data much. The Faculty of Medicine, then, is located at the University Medical Center Utrecht (UMCU) and is not of primary interest for the current survey, as it concerns an entirely different organization and research support structure.

```{r facultyplot}
# R color brewers palettes: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
# UU colors: https://www.uu.nl/organisatie/huisstijl/huisstijlelementen/kleur
UU_pallette <- c("#FFCD00", "#FFE6AB", "#F3965E", "#AA1555", "#6E3B23", "#24A793", "#5287C6", "#001240", "#5B2182")

# TODO: 
# Faculty_1 should be renamed into "Geosciences" -> mutate([nameofcontent] = Faculty_1)
# or newname = unlist(dppsurvey %>% select(Faculty_1) %>% filter(!is.na(Faculty_1)) %>% unique())
# rename(newname = oldname)
# Or make a dictionary variable?

# Simple barplot of faculty division (WORKS, yet to prettify)
dppsurvey %>% 
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty) %>%
  summarise(Count = length(Faculty)) %>%
  arrange(Count) %>% # Order by count
  mutate(Faculty = factor(Faculty, levels=Faculty)) %>% # Update factor levels
  ggplot(aes(x = Faculty, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Respondents per UU Faculty") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")
```

As can be seen in the plot below, most respondents are early career researchers (PhDs, junior researchers and postdoctoral researchers). We suspect that this is because 1) there simply are more early-career researchers than mid- and late career researchers, and 2) early-career researchers experience a greater need for support with respect to handling personal data.

```{r positionplot}
# Simple barplot of position division (WORKS, yet to prettify)
dppsurvey %>% 
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Position) %>%
  summarise(Count = length(Position)) %>%
  arrange(Count) %>% # Order by count
  mutate(Position = factor(Position, levels=Position)) %>% # Update factor levels
  ggplot(aes(x = Position, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Respondents per position type") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")
```

Below the division of position per faculty is plotted:

```{r facultypositionplot}
# Barplot of faculty and position (WORKS but looks bad and is NOT ordered)
dppsurvey %>%
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value = TRUE), 
               values_drop_na = TRUE, 
               values_to = "Faculty") %>%
  select(-name) %>%
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value = TRUE),  
               values_drop_na = TRUE, 
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Faculty, Position) %>%
  summarise(n = length(Faculty)) %>%
  ggplot(aes(x = Faculty, y = n, fill = Position)) + 
  geom_bar(stat = "identity") +
  labs(x = "", y = "Count", title = "Respondents per UU Faculty and Position") +
  scale_fill_manual(values = UU_pallette) + 
  coord_flip() +
  geom_text(aes(label = n), position= position_stack(vjust = 0.5)) + 
  theme_classic() + 
  theme(legend.position = "bottom")
  

## Test dataset - this works
#testdf <- data.frame(fac1=c("fsw",NA,NA,"fsw",NA), 
#                     fac2 = c(NA,"geo",NA,NA,NA),
#                     fac3 = c(NA,NA,"dgk",NA,"dgk"))

#testdf %>%
#  pivot_longer(1:3, values_drop_na = TRUE, values_to = "Faculty") %>%
#  select(-name) %>%
#  group_by(Faculty) %>%
#  summarise(n = length(Faculty)) %>%
#  ggplot(aes(x = Faculty, y = n)) + geom_bar(stat = "identity")
  
### 

### Testdataset 2 WORKS
#testdf2 <- data.frame(fac1 =c("fsw",NA,NA,"fsw",NA), 
#                     fac2 = c(NA,"geo",NA,NA,NA),
#                     fac3 = c(NA,NA,"dgk",NA,"dgk"),
#                     pos1 = c("PhD", NA, NA, "PhD",NA),
#                     pos2 = c("NA", "postdoc", "postdoc", NA, "postdoc"))

#testdf2 %>%
#  pivot_longer(contains("fac"), values_drop_na = TRUE, values_to = "Faculty") %>%
#  select(-name) %>%
#  pivot_longer(contains("pos"), values_drop_na = TRUE, values_to = "Position") %>%
#  select(-name) %>%
#  group_by(Faculty, Position) %>%
#  summarise(n = length(Faculty)) %>%
#  ggplot(aes(x = Faculty, y = n, fill = Position)) + geom_bar(stat = "identity")

``` 

## Types of data

```{r typesofdata}
# Link for grids for graphs: https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html

# Data types
# Simple barplot of types of data used (WORKS, yet to prettify)
datatypeplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  ggplot(aes(x = Datatype, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Data types") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Personal data types
personaldatatypeplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  ggplot(aes(x = Datatype, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Personal data types") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# From the gridExtra package
grid.arrange(datatypeplot, personaldatatypeplot, nrow = 1)

# Other data types not in these graphs
otherdatatypes <- 
  dppsurvey$Datatype_11_TEXT[!is.na(dppsurvey$Datatype_11_TEXT)] %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

#otherdatacounts <- as.data.frame(sort(table(otherdatatypes), decreasing=TRUE))
#otherdata <- wordcloud2(data = otherdatacounts, size=1.6, color=rep_len(UU_pallette, nrow(otherdatacounts)),
#            minRotation = 0, maxRotation = 0, rotateRatio = 1)

otherpersdatatypes <- 
  dppsurvey$Personaldata_type_10_TEXT[!is.na(dppsurvey$Personaldata_type_10_TEXT)] %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

#otherdatacounts <- as.data.frame(sort(table(otherdatatypes), decreasing=TRUE))
#otherpersdata <- wordcloud2(data = otherdatacounts, size=1.6, color=rep_len(UU_pallette, nrow(otherdatacounts)),
#            minRotation = 0, maxRotation = 0, rotateRatio = 1)

# grid.arrange(otherdata, otherpersdata, nrow = 1)

```

## Organisational and technical measures

### Organisational measures and storage media used

```{r orgmeasures}
# Link for grids for graphs: https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html

# Organisational measures
# Simple barplot of types of data used (WORKS, yet to prettify)
orgmeasuresplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Orgmeasures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Orgmeasures") %>%
  select(-name) %>%
  group_by(Orgmeasures) %>%
  summarise(Count = length(Orgmeasures)) %>%
  arrange(Count) %>% # Order by count
  mutate(Orgmeasures = factor(Orgmeasures, levels=Orgmeasures)) %>% # Update factor levels
  ggplot(aes(x = Orgmeasures, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Organisational measures") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

#orgmeasuresplot

# Other organisational measures
otherorgmeasures <- 
  dppsurvey$Orgmeasures_16_TEXT[!is.na(dppsurvey$Orgmeasures_16_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

# Storage media used
storagemediaplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Storage_medium_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Storagemed") %>%
  select(-name) %>%
  group_by(Storagemed) %>%
  summarise(Count = length(Storagemed)) %>%
  arrange(Count) %>% # Order by count
  mutate(Storagemed = factor(Storagemed, levels=Storagemed)) %>% # Update factor levels
  ggplot(aes(x = Storagemed, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Storage media used") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

#storagemediaplot

# Other storage media
otherstoragemedia <- 
  dppsurvey$Storage_medium_8_TEXT[!is.na(dppsurvey$Storage_medium_8_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

grid.arrange(orgmeasuresplot, storagemediaplot, nrow = 1)
```

### Informed consent forms

```{r consent}
# Do you use consent forms in your research?
consentformplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consentforms_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "consentforms") %>%
  select(-name) %>%
  group_by(consentforms) %>%
  summarise(Count = length(consentforms)) %>%
  arrange(Count) %>% # Order by count
  mutate(consentforms = factor(consentforms, levels=consentforms)) %>% # Update factor levels
  ggplot(aes(x = consentforms, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Do you use consent forms?") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Consent forms: other
consentformsother <- 
  dppsurvey$Consentforms_7_TEXT[!is.na(dppsurvey$Consentforms_7_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

# Consent form content
consentcontentplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consent_content_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "consentconsent") %>%
  select(-name) %>%
  group_by(consentconsent) %>%
  summarise(Count = length(consentconsent)) %>%
  arrange(Count) %>% # Order by count
  mutate(consentconsent = factor(consentconsent, levels=consentconsent)) %>% # Update factor levels
  ggplot(aes(x = consentconsent, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Content of the consent form") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Consent form content: other
consentcontentother <- 
  dppsurvey$Consent_content_12_TEXT[!is.na(dppsurvey$Consent_content_12_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

grid.arrange(consentformplot, consentcontentplot, nrow = 1)
```

### Data Protection Impact Assessment (DPIA)

"DPIA_Help"


```{r dpia}
# Experience with DPIAs?
dpiaexperienceplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_experience_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "dpiaexperience") %>%
  select(-name) %>%
  group_by(dpiaexperience) %>%
  summarise(Count = length(dpiaexperience)) %>%
  arrange(Count) %>% # Order by count
  mutate(dpiaexperience = factor(dpiaexperience, levels=dpiaexperience)) %>% # Update factor levels
  ggplot(aes(x = dpiaexperience, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Experience with DPIAs") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# DPIA: had help?
dpiahelpplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_Help_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "dpiahelp") %>%
  select(-name) %>%
  group_by(dpiahelp) %>%
  summarise(Count = length(dpiahelp)) %>%
  arrange(Count) %>% # Order by count
  mutate(dpiahelp = factor(dpiahelp, levels=dpiahelp)) %>% # Update factor levels
  ggplot(aes(x = dpiahelp, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Help with DPIA and from whom") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# DPIA help from: other
dpiahelpother <- 
  dppsurvey$DPIA_Help_4_TEXT[!is.na(dppsurvey$DPIA_Help_4_TEXT)] %>%
  unlist() %>%
  str_trim() %>%
  str_to_lower()

grid.arrange(dpiaexperienceplot, dpiahelpplot, nrow = 1)
```

## Data sharing and publication
Share_outside_UU", "Share_measures",
Data_publication", "Publication_format", "Reason_NoPub"


## Existing support

"Know_PO", "Lookedforhelp", "Foundhelp",
Sourcesused", 

## Support required

Obstacle, Challenges, Infotools_missing, Beter_support

### Interviews

There are currently `r length(dppsurvey$Email_1_TEXT[!is.na(dppsurvey$Email_1_TEXT)])` researchers who have left their email address to be contacted. Below you can find from which faculties they are:

```{r demo-emailpeople}
# Select the people that left their email address, and their
# Faculty, department and position (combined in 1 column)
emaildataset <- dppsurvey %>%
  filter(!is.na(Email_1_TEXT)) %>%
  select(contains(c("Faculty", "Dept", "Position", "Email_1"))) %>%
  rename(Email = Email_1_TEXT) %>%
  unite("Faculty", grep("Faculty",names(.), value = T), 
        sep = ",", remove = TRUE, na.rm=TRUE) %>%
  unite("Dept", grep("Dept",names(.), value = T), 
        sep = ",", remove = TRUE, na.rm=TRUE) %>%
  unite("Position", grep("Position",names(.), value = T), 
        sep = ",", remove = TRUE, na.rm=TRUE)

# Plot the division of email addresses over faculties and positions
UU_pallette <- c("#FFCD00", "#FFE6AB", "#F3965E", "#AA1555", 
                 "#6E3B23", "#24A793", "#5287C6", "#001240", "#5B2182")

emaildataset %>% 
  ggplot(aes(x = Faculty, fill = Position)) +
  geom_bar(stat = "count") + 
  stat_count(geom = "text", aes(label = ..count..), position = position_stack(vjust=0.5)) + 
  coord_flip() + 
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = UU_pallette) +
  labs(x = "", y = "Count", title = "Email people per faculty and position")

# For later: looping through variables to combine them
#variables <- c("Faculty", "Dept", "Position", "Datatype", 
#               "Personaldata_type", "Orgmeasures","Storage_medium",
#               "ConsentForms", "Consent_content", "DPIA_experience", 
#               "DPIA_Help", "Share_outside_UU", "Share_measures",
#               "Data_publication", "Publication_format", "Reason_NoPub",
#               "Know_PO", "Lookedforhelp", "Foundhelp",
#               "Sourcesused", "Obstacle", "Challenges",
#               "Infotools_missing", "Better_support",
#               "Email", "Email_1", "Duration")

#for(varname in variables){ # Does not work yet
#  emailpeople2 <- emailpeople %>%
#    unite(varname, grep(varname, names(emailpeople), value = TRUE), sep = ",", remove = TRUE, na.rm=TRUE)
#}

todaysdate <- as.character(Sys.Date())
write.csv(emaildataset, file.path(paste0("../../", todaysdate, "_emailaddresses.csv")))

```

## OLD
### Open text responses

There are many questions with open text responses (questions with an "Other, namely" option and Infotools_missing and Better_support variables), which are harder to analyze but extremely important to take into account in the Data Privacy Project

```{r selectopenquestions}
#opentextresponses <- dppsurvey %>% 
#  select(Faculty, Position, Challenges, Infotools_missing, Better_support_9_TEXT)
#
#write.csv(opentextresponses, file.path(paste0("../data/raw/", todaysdate, "_dppsurvey_opentextresponses.csv")))
```


## Technical info

This document was last created on: `r as.character(Sys.Date())`.

```{r sessioninfo}
sessionInfo()
```