---
title: "Data Privacy Survey"
subtitle: "Recommendations for improving privacy-related services for research at Utrecht University"
author:
- "Dorien Huijser, Research Data Management Support, Utrecht University"
- "Rik Janssen, Information Technology Services, Utrecht University"
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
abstract: |
  In the second quarter of 2022, Utrecht University (UU) Research Data Management 
  Support (RDM Support) sent out a survey among all scientific personnel at UU: 
  the Data Privacy Survey. The aim of this survey was to investigate 1) How UU 
  researchers currently deal with personal data in research, 2) What challenges 
  they run into when handling personal data in research, and 3) How support at
  UU can improve their services concerning personal data in research. The 
  results show that most researchers know to take into account 
  privacy in their projects. However, there are vast differences in knowledge 
  on this topic, as well as confusion on topics such as data sharing, 
  anonymisation, and the administrative requirements surrounding privacy. Taken
  together, the survey shows that researchers could benefit from research support
  that is more visible in the organisation and more hands-on (i.e., advise that is
  immediately applicable instead of more general). Additionally, more awareness 
  of the most important privacy-related topics, and how to deal with them in 
  research seems desirable, both for newly employed and existing researchers.
  Finally, many researchers indicate to experience a high administrative burden 
  that is partly caused by privacy-related documentation before starting a 
  project. Finding a way to decrease this burden while still complying with 
  legal and policy regulations will likely help to increase compliance and 
  help researchers focus on performing high-quality research.
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
    df_print: paged
    css: "../assets/style.css"
    includes:
      in_header: "../includes/header.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

```{r sourcescript, include=FALSE, cache=FALSE}
knitr::read_chunk('plot-data.R')
```

# Introduction {#introduction}
Scientific research often includes some form of personal data. The amount of 
personal data that is processed in research is likely underestimated, as 
researchers may be unaware of what personal data are or whether they are being 
collected. With the implementation of the General Data Protection Regulation
(GDPR) in 2018, stricter legal requirements apply to handling personal data and 
its sharing and publication. In our own experience, the number and complexity of 
questions on handling personal data in scientific research at Utrecht University
(UU) is increasing. 

Our premise at <a href = "https://uu.nl/rdm" target = "_blank">Research Data 
Management Support</a> (RDM Support) is to assist researchers with any issues 
surrounding the management of their research data, including research data that 
contains personal data. To understand how we can best help researchers with 
their privacy-related questions and needs, we wanted to investigate:

1. To what extent are UU researchers aware of privacy legislation and practices?
2. What data privacy issues do UU researchers typically run into?
3. What support do UU researchers need to handle personal data?

To answer these questions, we set up an online survey and planned one-on-one 
meetings with a selection of UU researchers. This report summarises our findings,
and describes recommendations to improve privacy-related services for research
at Utrecht University. For a full description of the methods and results, please 
refer to the
<a href = "data-privacy-survey-report" target = "_blank">Data Privacy Survey Report</a>.

This survey was part of a larger project at UU, the <a href = "https://utrechtuniversity.github.io/dataprivacyproject/" 
target = "_blank">Data Privacy Project</a>^[The Data Privacy Project is funded 
by Utrecht University’s Research IT program and a Digital Competence Center 
grant from the Dutch Organization for Scientific Research (NWO). It is led by Utrecht 
University's <a href="https://uu.nl/rdm/" target = "_blank">Research Data 
Management Support</a>, in collaboration with the University Library, 
Information Technology Services, first- and second-line privacy officers, 
information security, and others.]. This project is a data support effort within 
UU that aims to provide actionable and FAIR (Findable, Accessible, Interoperable, 
Reusable) information and tools for researchers to handle personal data in their 
research. 

# Methods summary {#methods}
```{r dependencies}
```

```{r readdata}
```

The Data Privacy Survey consisted of two main parts:

- An online survey that was sent to all of UU's scientific personnel via email, 
asking about researchers' common data types, current practices
in handling personal data in their research, awareness of data support channels, 
and challenges and needs in this area (n = `r dim(dppsurvey)[1]`).
- One-on-one meetings with individual researchers to talk about their experiences,
challenges and needs in this area (n = 27).

All relevant materials (including this report) can be found in the 
<a href = "https://github.com/utrechtuniversity/dataprivacysurvey" target = "_blank">GitHub repository</a>.

# Results summary {#results-summary}
In short, we found that:

- Personal data are processed in each faculty and across all academic positions
at Utrecht University, but mostly among the faculties of Science and Social and
Behavioural Sciences, and among early-career researchers.
- Most researchers deal with tabular, textual, code, and audiovisual data, which
contain mainly demographics, contact information, and direct identifiers.
- Most researchers seem to be at least familiar with concepts like anonymisation 
and pseudonymisation, access control, and UU-approved tools for handling personal 
data. However, the knowledge level differs quite a bit. For example:
  - some researchers seem to store data on locations that officially are not allowed.
  - information to study participants is sometimes lacking all the required information, 
  such as contact information from the Data Protection Officer, and information and
  consent for sharing data for reuse purposes.
  - about half of the sample is not familiar with the concept of a Data 
  Protection Impact Assessment.
  - data are often shared, but not all the right measures are always taken to 
  do so securely. 
- Privacy is sometimes found to be an obstacle for open science, also confirmed by 
a relatively low amount of published datasets.
- Not all help is found equally well. For example, about half of the respondents
indicated not to be aware of their faculty privacy officer, and many researchers
resort to the UU website, intranet, and colleagues for help.
- Researchers may benefit from greater visibility and findability of support, 
specific tools. more and more hands-on support, more clear and less bureaucratic
processes and guidelines, and specific information and education.

Together, these findings suggest that there is still a lot to be gained
regarding awareness and secure practices in scientific research with personal data.

## Recommendations {#recommendations}
Logische flow: 

- Teveel werk, we moeten het minder belastend maken
- Information provision
  - op teveel plekken, 
  - te weinig bruikbare informatie: concrete informatie bijv. templates (data processing agreements, DPIA) en tools. User 
     story: gaan gewoon zelf googlen en komen dan niet op de goeie uit. Daarna weer 
     veel tijd kwijt om dit recht te trekken maar soms te laat.
  - wat wordt aangeraden door UU
  - inhoudelijk: op tijd beginnen, 
- Meedenken/hands-on support. Meer denken vanuit onderzoek, meer idee hebben waar onderzoek over gaat, wat het betekent om onderzoek te doen, waar onderzoekers tegenaan lopen en wat haalbare oplossingen zijn. 
- Administratie rondom onderzoek is een chaos en teveel en teveel verschillen tussen faculteiten


---
Awareness: start early

Information findability
Doorzoekbaarheid huidige pagina's, cluster informatie op 1 plek, verwijs naar 
eigen pagina's. 

Als ze info vinden: hebben ze er dan wat aan?
Wekle informatie nodig
Duurt te lang: meer staff of betere informatie

Aan de slag met privacy in onderzoek: mag het wat minder admin



# OLD STUFF
### Increased visibility and findability {#increased-visibility-and-findability}
The findings show that the visibility of existing support for handling personal 
data can be increased at UU. This includes visibility of privacy officers, 
legal affairs and Research Data Management Support, as well as faculty- or 
department-level support. 

As shown by the fact that many respondents consult with colleagues for 
information and support, personal contact with supporters will likely be more 
effective at increasing awareness and uptake of secure privacy practices than a 
passive website that may not always answer all possible questions.

Some suggestions for increasing visibility and findability:

- Make sure there is (sufficient) support available to help researchers in time.
If there is no support available, the odds are that researchers may not ask for 
help at all and thus forget important privacy elements in the design of their 
research project.
- Make sure support staff are easy to find. For example, give them a clear location
on the university website and/or on the intranet. Also indicate who is available
for what type of support and for which target audience (e.g., "all UU 
researchers", "employees from the Faculty of Humanities", etc.)
- Make sure information on handling personal data in research is easy to find.
The UU website and intranet seem the most suitable places, as these are the most
visited to look for information according to the survey. 
- Make researchers actively aware of existing support channels. For example,
first-line support can visit important faculty or department events, offer to
give a workshop, and keep short lines of contact with department heads.

### More hands-on support {#more-hands-on-support}
In line with the [findings described above](#hands-on-support), more hands-on 
support is desirable. This can for example be reached by:

- More collaboration between different types of data support. For example, at UU
there is a clear distinction between privacy officers (employed by faculties),
data managers (employed by the University Library and some faculties), information
security officers (Information Technology Services) and research engineers 
(Information Technology Services). However, in practice there is often overlap 
between either or all of these positions. Regular collaboration between these 
parties can:
  - help privacy and security experts understand the perspective of the 
  researcher and possibly how to apply their advice in practice,
  - help data managers and research engineers understand the GDPR requirements 
  and translate those to practice, and
  - enable reuse of solutions already applied in earlier cases that either 
  party may not be aware of.
- Together with researchers, developing suitable practical solutions (tools, 
protocols, etc.) for specific scenarios and/or types of research, and sharing 
them with the world for reuse (if possible).
- Increasing contact between support staff and researchers to understand the 
issues they are dealing with that require a solution.

### Decreased administrative burden {#decreased-administrative-burden}

In line with the [findings described above](#less-bureaucracy), it is desirable
to attempt to decrease the administrative burden on researchers. We suggest that 
the bureaucracy issues can be partly attributed to:

1. Understaffing: it is currently difficult for data support staff to review 
*all* research projects involving personal data at UU. This can lead to 
projects with issues being noticed too late and researchers having 
to make changes to their project after the fact. This creates unnecessary 
administrative burden that could have been prevented had researchers been 
in contact with support staff from early on. 
2. Lack of unifying policy on steps to take when working with personal data: At
the moment each faculty has its own procedure, or has none at all. This can 
create confusion among both researchers and support staff on which steps to take
when, who to contact for help, or who is responsible for deciding on compliance.
3. Lack of integration between systems and staff: having to fill out the same
thing multiple times may be caused by the fact that different administrative
systems are used for different purposes. Integration of these systems may lead
to a relief of the administrative burden experienced by researchers. <br>For 
example, DMPonline is used at UU for registering Data Management Plans. However, 
DMP templates do not include sufficient questions about how personal data are 
handled, and thus a privacy scan may be additionally required. Privacy is 
sometimes part of an ethical application, and so this same information may be 
included there again as well. Finally, the processing register also requires 
information on (among others) the type and sensitivity of personal data, 
which is a third system where this information will be registered.

### Education and outreach {#education-and-outreach}

Many problems seem to arise from the fact that privacy has not been taken into
account from the design phase of a research project. This leads to research
support staff being contacted too late in the process, who then has to tell the
researcher to retrospectively make changes in their design in order to move 
forward. Such scenarios may be prevented by, for example:

- Including the basic of handling personal data and contact details of support
staff in onboarding processes when researchers are newly hired.
- Educating researchers on taking privacy into account in the designing phase
of their project. 
- Organising outreach events, such as the "Privacy and Security week" making
researchers aware of privacy-related measures (see further Visibility).
- Embedding privacy in existing materials and processes, such as in ethical
applications, in the education of research master and PhD students, etc.

### Specific information and tools {#specific-info-tools}
The specific [information](#faq) and 
[tools](specific-tools) described above can be addressed in current and future
support efforts. One of these is the 
<a href="https://utrechtuniversity.github.io/dataprivacyproject">Data Privacy Project</a>, 
from which this Data Privacy Survey forms a part. Other efforts may include the
updating of websites from RDM Support, faculties, or ethical committees, and the
updating of services offered by the IT department.
