---
title: "Data Privacy Survey"
subtitle: "Challenges and needs of privacy-related services for research at Utrecht University"
author:
- "Dorien Huijser"
- "Neha Moopen"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
  In the second quarter of 2022, Utrecht University (UU) Research Data Management 
  Support (RDM Support) sent out a survey among all scientific personnel at UU: 
  the Data Privacy Survey. The aim of this survey was to investigate 1) How UU 
  researchers currently deal with personal data in research, 2) What challenges 
  they run into when handling personal data in research, and 3) How support at
  UU can improve their services concerning personal data in research. The 
  results show that most researchers know to take into account 
  privacy in their projects. However, there are vast differences in knowledge 
  on this topic, as well as confusion on topics such as data sharing, 
  anonymisation, and the administrative requirements surrounding privacy. Taken
  together, the survey shows that researchers could benefit from research support
  that is more visible in the organisation and more hands-on (i.e., advise that is
  immediately applicable instead of more general). Additionally, more awareness 
  of the most important privacy-related topics, and how to deal with them in 
  research seems desirable, both for newly employed and existing researchers.
  Finally, many researchers indicate to experience a high administrative burden 
  that is partly caused by privacy-related documentation before starting a 
  project. Finding a way to decrease this burden while still complying with 
  legal and policy regulations will likely help to increase compliance and 
  help researchers focus on performing high-quality research.
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
    df_print: paged
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Introduction
Scientific research often includes some form of personal data. The amount of 
personal data that is processed in research is likely underestimated, as 
researchers may be unaware of what personal data are or whether they are being 
collected. With the implementation of the General Data Protection Regulation
(GDPR) in 2018, stricter legal requirements apply to handling personal data and 
its sharing and publication. In our own experience, the number and complexity of 
questions on handling personal data in scientific research at Utrecht University
(UU) is increasing. 

Our premise at <a href = "https://uu.nl/rdm" target = "_blank">Research Data 
Management Support</a> (RDM Support) is to assist researchers with any issues 
surrounding the management of their research data, including research data that 
contains personal data. To understand how we can best help researchers with 
their privacy-related questions and needs, we wanted to investigate:

1. To what extent are UU researchers aware of privacy legislation and practices?
2. What data privacy issues do UU researchers typically run into?
3. What support do UU researchers need to handle personal data?

To answer these questions, we set up an online survey and planned one-on-one 
meetings with a selection of UU researchers. This report describes our methods, 
results and the future steps that data support at UU can take to improve their 
services concerning privacy in research.

This survey was part of a larger project at UU, the <a href = "https://utrechtuniversity.github.io/dataprivacyproject/" 
target = "_blank">Data Privacy Project</a>^[The Data Privacy Project is funded 
by Utrecht University’s Research IT program and a Digital Competence Center 
grant from the Dutch Organization for Scientific Research (NWO). It is led by Utrecht 
University's <a href="https://uu.nl/rdm/" target = "_blank">Research Data 
Management Support</a>, in collaboration with the University Library, 
Information Technology Services, first- and second-line privacy officers, 
information security, and others.]. This project is a data support effort within 
UU that aims to provide actionable and FAIR (Findable, Accessible, Interoperable, 
Reusable) information and tools for researchers to handle personal data in their 
research. 

# Methods
```{r dependencies}
# install.packages("data.table")
library(data.table)

# install.packages("tidyverse")
library(tidyverse)

#install.packages("gridExtra")
library(gridExtra)

#install.packages("knitr")
library(knitr)

#install.packages("kableExtra")
library(kableExtra)
```

```{r readdata}
### Find the most recent data file in data/pseud or data/processed and read it in
# List all the files in pseud and processed folder
data_files_pseud <- list.files(path = "../data/pseud")
data_files_processed <- list.files(path = "../data/processed")

data_files <- data.frame(filename = c(data_files_pseud,
                                      data_files_processed),
                         folder = c(rep("pseud", length(data_files_pseud)),
                                    rep("processed", length(data_files_processed))
                                    )
                         )

# Get the dates out of the file names and pick the newest one
data_files$filenamedates <- as.Date(str_extract(pattern = "[0-9]+-[0-9]+-[0-9]+", 
                                     string = data_files$filename),
                         format = "%Y-%m-%d")

# Sort by date (descending = most recent one first)
setorder(data_files, -filenamedates, na.last = TRUE)

# Get most recent file
datafile <- paste0("../data/",
                   data_files$folder[1],
                   "/",
                   data_files$filename[1])

dppsurvey <- fread(datafile)
```

## Survey questions 
The [Data Privacy Survey](#data-and-material-availability) was created by the project 
coordinators (DH, NM) with input from a wide variety of experts, consisting of 
privacy experts, data managers, data consultants, and IT staff (e.g., information 
security,  research engineering). The survey consisted of 19-23 questions 
(the exact number differing depending on given answers), which were separated 
into 5 sections:

1. **About you and your research**: 5 questions about the faculty, department, 
position, and type of (personal) data the researcher works with.
2. **Measures and documents**: 4-6 questions about researchers' familiarity and
use of privacy-related measures (e.g., processing register, Data Privacy Impact
Assessment, encryption, pseudonymisation, etc.), data storage, and informed 
consent forms.
3. **Data sharing, archiving and publication**: 3-4 questions about researchers' 
data sharing and publishing practices.
4. **Finding support**: 3-4 questions about the awareness of several data support
channels.
5. **Improving our services**: 4 questions about researchers' challenges, needs, 
and suggestions to improve UU's data privacy support.

The estimated 10 minute duration of the survey in advance proved to be 
relatively accurate: respondents spent a median of 
`r round(median(dppsurvey$Duration/60), 1)` minutes on the survey (although 
there was a very large variation; range: 
`r round(min(dppsurvey$Duration)/60, 1)` minutes - 
`r round(max(dppsurvey$Duration)/3600/24,1)` days).

## Procedure online survey
The survey was created in Qualtrics and distributed from March 21st, 2022 
onwards via several communication channels to reach as many UU researchers as 
possible:

- An <a href = "https://utrechtuniversity.github.io/dataprivacyproject/20220321-data-privacy-survey.html"
target = "_blank">email</a> was sent to all academic staff at UU through 
central communication channels. The reasoning was that this would be the most
effective way to reach as many UU researchers as possible - taking for granted
that we would likely miss a small part of non-academic personnel also involved
in research in some way.
- A mention in several Faculty newsletters.
- Social media messages (e.g., on <a href = "https://twitter.com/RDMsupportUU/status/1517390979707412481"
target = "_blank">Twitter</a>).
- A <a href = "https://intranet.uu.nl/en/news/news-items/survey-better-support-for-handling-privacy-sensitive-data-in-research" target = "_blank">news item</a> on the UU intranet.
- Via data support colleagues, who were asked to point researchers they were
in contact with to the survey.

Researchers could decide voluntarily whether they wanted to participate in the
survey. All results reported here are from researchers who provided their active 
consent, worked at Utrecht University and indicated to work with some type of 
research data.

## Online survey respondents
The survey was filled out by `r dim(dppsurvey)[1]` UU researchers. As can be 
seen in the figure below, we received responses from each UU faculty, but the 
distribution was not equal: the faculties of Science and Social and Behavioural 
Sciences were overrepresented, whereas the responses of the Faculty of 
Geosciences and the Faculty of Medicine were rather low. This can be explained 
by the mere size of the faculties (e.g., the faculty of Science is UU's largest 
faculty), but also by the types of research performed there. For example, 
research performed at the faculty of Geosciences is largely involved with 
natural scientific data, rather than data from or relating to humans. The 
Faculty of Medicine, then, is located at the University Medical Center Utrecht 
(UMCU) and is not of primary interest for the current survey, which was sent to 
UU researchers only. A small selection of respondents indicated to work at 
another part of the organisation, being mainly University College Utrecht and 
the University Corporate Office.

Most of the survey respondents were relatively early 
career researchers (PhDs, junior researchers and postdoctoral researchers). 
We suspect that this is because 1) there simply are more people in these positions
than there are in the others, and 2) early-career researchers may experience a 
greater need for support with respect to handling personal data.

```{r facultyplot}
# R color brewers palettes: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
# UU colors: https://www.uu.nl/en/organisation/corporate-identity/brand-policy/colour
UU_pallette <- c(#"#FFCD00", # UU yellow
                 "#FFE6AB", # Lighter yellow
                 "#F3965E", # Orange
                 "#AA1555", # Bordeaux-red
                 "#6E3B23", # Brown
                 "#24A793", # Green
                 "#5287C6", # Blue
                 "#001240", # Dark blue
                 "#5B2182") # Purple

# Plot of positions per faculty
position_per_faculty <- dppsurvey %>%
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Faculty, Position) %>%
  summarise(Count = length(Faculty)) %>%
  ggplot(aes(reorder(Faculty, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Position)) +
  coord_flip() +
  theme_classic() + 
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  labs(x = "", 
       title = "Faculty and Position representation",
       caption = "Survey respondents per faculty and position") +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  #scale_x_continuous(expand = expansion(add = 3)) + 
  geom_label(aes(label = after_stat(y), group = Faculty), # totals per faculty
             stat = 'summary', 
             fun = sum, 
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             nudge_y = 2,
             label.size = NA, #1,
             label.padding = unit(0.35, "lines"),
             #label.r = unit(0.25, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  geom_text(aes(label = Count), # counts per position per faculty
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))

position_per_faculty

```

## One-on-one meetings
Besides the online survey, we organised one-on-one meetings with researchers, to 
hear about their personal experiences, challenges and needs concerning the handling 
of personal data in their research. Survey respondents could voluntarily leave 
their email address at the end of the survey to be contacted by us. These 
meetings were semi-structured and revolved around the following questions:

- What made you leave your email address in the Data Privacy Survey? Related:
What are your general experiences in handling personal data in research?
- Which difficulties do you run into when handling personal data?
- What support would you need to help you handle personal data in your research?
- Do you have a concrete need for support at the moment?

 From the survey respondents, 
`r sum(dppsurvey$Email == "Yes", na.rm=TRUE)` researchers left their email 
address to be contacted. Of those, 28 researchers indicated that they were 
willing to meet with us. So far, 27 one-on-one online 
meetings have been conducted, each taking about 30 minutes. During the meetings,
one of the project coordinators (DH) was always present, together with either the
other project coordinator (NM) or the relevant faculty privacy officer. Before
the privacy officer was invited to the meeting, the researcher's consent to do 
so was always asked first.

Below, the division over faculties can be seen for all interviewees. For privacy
reasons, the positions of the interviewees are not displayed here. 

```{r emailaddressplot}
library(readxl)

# Read in coded responses from interviews
interviewfile <- "../data/coded/dppsurvey_interviews_coded.xlsx"
interviews <- read_excel(interviewfile, sheet = "interviews")

meetings_faculties <- data.frame(Faculty = str_trim(unlist(str_split(interviews$Faculty, 
                                                                     ",")), 
                                                    side = "both"))

emailsplot <-
  meetings_faculties %>%
  group_by(Faculty) %>%
  summarise(Count = length(Faculty)) %>%
  arrange(Count) %>% # Order by count
  mutate(Faculty = factor(Faculty, levels=Faculty)) %>% # Update factor levels
  ggplot(aes(x = Faculty, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  labs(x = "", 
       title = "Faculty representation one-on-one meetings",
       caption = "Faculty representation of researchers spoken to in one-on-one meetings") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             label.size = NA, #1,
             label.padding = unit(0.35, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  theme(axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank())

emailsplot

# Code for donut chart - looks nice, but is horrible to actually interpret
#donutchart <-
#  meetings_faculties %>%
#  group_by(Faculty) %>%
#  summarise(Count = length(Faculty)) %>%
#  arrange(Count) %>% # Order by count
#  mutate(Faculty = factor(Faculty, levels=Faculty)) %>% # Update factor levels
#  mutate(Fraction = (Count/sum(Count))) %>%
#  mutate(ymax = cumsum(Fraction)) %>%
#  mutate(ymin = c(0, head(ymax, n = -1))) %>%
#  mutate(labelPosition = (ymax + ymin)/2) %>%
#  mutate(label = paste0(Count)) %>%
#  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = Faculty)) +
#  geom_rect() +
#  geom_label(x = 3.5,
#             aes(y = labelPosition, label = label),
#             size = 3,
#             fill = "white") + 
#  coord_polar(theta = "y") +
#  xlim(c(1, 4)) + 
#  scale_fill_manual(values = UU_pallette) +
#  theme_void()
#
#donutchart
```

## Analysis
From the raw Qualtrics output, we first cleaned and split the data into different
data files (cleaned and closed survey responses, open text responses, email
addresses, see the 
<a href="https://github.com/UtrechtUniversity/dataprivacysurvey/blob/main/src/pseudonymise-data.R" 
target="_blank">pseudonymise-data.R</a> script for details). Both the open text responses and 
the notes made during the one-on-one meetings were separately and manually scored
to enable the extraction of action points. 

Below we report on the descriptive statistics or summaries from most of the 
survey questions. As we did not formulate hypotheses, no statistical analyses 
were performed. 

## Data and material availability
All survey-related documentation can be found in the <a href="https://github.com/UtrechtUniversity/dataprivacysurvey" 
target = "_blank">dedicated survey repository</a>. It contains:

- The 
<a href = "https://utrechtuniversity.github.io/dataprivacysurvey/assets/survey-questions-qualtrics.pdf"
target = "_blank">full survey</a>
- The survey's <a href = "https://utrechtuniversity.github.io/dataprivacysurvey/assets/survey-privacy-statement.pdf"target = "_blank">privacy statement</a>
- The <a href="https://github.com/UtrechtUniversity/dataprivacysurvey/tree/main/src" target = "_blank">
scripts used to clean the data and write this report</a>

As the dataset contains personal information (demographic information, open text 
responses, email addresses, etc.), and no consent was obtained to share those 
details, we are unable to share them in this repository.

# Results
## Types of research data {.tabset}
### University-wide

To investigate the types of research that was represented in the sample, we 
asked which types of data, and specifically which types of *personal* data the 
respondents worked with. Most respondents indicated to use tabular, textual, 
code and audio data as their primary research data format. 
In terms of *personal* data types, contact information, demographic information 
and direct identifiers were most common among the respondents. This 
can be either because these are the types of personal data that are indeed most 
common, but possibly also because researchers mostly recognise these types of 
data as being personal data.

```{r datatypes, results = "asis"}
# Table of data types
datatypetable <- dppsurvey %>% 
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                           " (",
                           Percentage,
                           "%)"),
         ) %>%
  select(-Count, -Percentage)

# Table of personal data types
tablepersdata <- dppsurvey %>% 
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                           " (",
                           Percentage,
                           "%)"),
         ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```


When comparing faculties, it appears that most personal data is processed in the
faculties of Social and Behavioural Sciences (FSBS), although the faculties of 
Science and Veterinary Medicine also seem to process quite some personal data.
In the tabs, we look further into the types of (personal) data processed within
each faculty.

```{r personaldataplot}
# Graph
datatypegraphs <- dppsurvey %>% 
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  pivot_longer(cols = grep("^Faculty_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  arrange(Percentage) %>% # Order by percentage
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Faculty, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", 
       title = "Personal data types across Faculties (%)",
       caption = "Percentage of the total sample size of personal data used in every UU faculty") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = paste0(Percentage, "%")),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

datatypegraphs

```



### Science
As with the university-wide data, researchers from the Science faculty indicated
that they also most often used tabular data, textual data and code/theoretical 
models in their research. The same goes for the types of *personal* data: the
same top-3 types are used here as in the entire university (demographic 
information, contact information, direct identifiers). 

```{r beta-datatypes, results = "asis"}
# Table of data types
beta_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_5)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
beta_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_5)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

beta_tablepersdata <- beta_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(beta_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(beta_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(beta_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(beta_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

As a large part of the researchers from the Science faculty indicate not to use
*any* personal data whatsoever in their research, it may be useful to 
investigate in which departments these researchers work. In the figure below,
the division of personal data types per Science department is plotted. As can be
expected, the departments of Biology, Physics and Mathematics do not seem to 
process much personal data, whereas in the Information and Computing Sciences
and Pharmaceutal Sciences departments, the most personal data is processed. This
makes sense considering the types of research performed in these departments.

```{r beta-personaldataplot}

# Graph
beta_datatypegraphs <- beta_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_Science_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across Science departments") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

beta_datatypegraphs
```

### FSBS
As with the university-wide data, researchers from the faculty of Social and
Behavioural Sciences (FSBS) indicated that they also most often used tabular data, 
textual data and code/theoretical models in their research. The same goes for 
the types of *personal* data: demographic and contact information are most often
used. Interestingly, sensitive demographic information is used much more relative
to  the UU-wide responses.

```{r fsw-datatypes, results = "asis"}
# Table of data types
fsw_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_6)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
fsw_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_6)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

fsw_tablepersdata <- fsw_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(fsw_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(fsw_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(fsw_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(fsw_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

When comparing the types of personal data across departments of the faculty, the
departments of Sociology and Cultural Anthropology seem to process the least
amount of personal data - although these departments may also simply be 
underrepresented in the survey sample. Most other departments process quite some
personal data, as can be expected from a faculty of which the research focuses
almost exclusively on humans. 

```{r fsw-personaldataplot}
# Graph
fsw_datatypegraphs <- fsw_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_FSW_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, nrow = 3,
             labeller = label_wrap_gen(width=11)) +
  labs(x = "", title = "Personal data types across FSBS departments") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

fsw_datatypegraphs
```

### Humanities
As with the university-wide data, researchers from the faculty of Humanities 
indicated that they also most frequently used tabular data, textual data and 
audio data in their research. The same goes for the types of *personal* data: the
same top-3 types are present here as in the entire university (demographic 
information, contact information, direct identifiers). 

```{r hum-datatypes, results = "asis"}
# Table of data types
hum_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_2)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
hum_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_2)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

hum_tablepersdata <- hum_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(hum_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(hum_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(hum_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(hum_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

If we look at the different departments within the faculty, it is clear that
mostly the departments of Languages, Literature and Communication, and of Media
and Culture Studies process personal data in their research. Especially 
researchers at the department of Languages, Literature and Communication seem to
process a lot of personal data, which makes sense considering the type of research
performed there.

```{r hum-personaldataplot}
# Graph
hum_datatypegraphs <- hum_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_Humanities_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 3,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across Humanities departments") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

hum_datatypegraphs
```

### Veterinary Medicine

When comparing the types of data that researchers at the Faculty of Veterinary
Medicine work with, it is clear that researchers at this faculty work with 
slightly different types of data, i.e. biological and physiological data are 
more common among these researchers. Despite this, the most frequent types
of personal data do correspond well to  the university-wide numbers: 
demographic information, direct identifiers and contact information are also 
the most common types of personal data researchers at this faculty seem to 
deal with.

```{r dgk-datatypes, results = "asis"}
# Table of data types
dgk_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_7)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
dgk_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_7)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

dgk_tablepersdata <- dgk_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(dgk_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(dgk_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(dgk_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(dgk_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

When looking at each department separately, it is clear that researchers in the
department of Population Health Sciences process the large majority of the 
personal data within the faculty. Of course, it may also be possible that this 
department was simply more represented in the survey than the other departments.

```{r dgk-personaldataplot}
# Graph
dgk_datatypegraphs <- dgk_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_DGK_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 3,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across Veterinary Medicine") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

dgk_datatypegraphs
```

### LEG
As with the university-wide data, researchers from the faculty of Law, 
Economics and Governance (LEG) indicated that they also most frequently used 
tabular data, textual data and audio data in their research. The same goes for 
the types of *personal* data: the same top-3 types are present here as in the 
entire university (demographic information, contact information, 
direct identifiers). 

```{r leg-datatypes, results = "asis"}
# Table of data types
leg_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_3)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
leg_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_3)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

leg_tablepersdata <- leg_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(leg_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(leg_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(leg_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(leg_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

When looking at the different departments within the faculty, researchers in 
the department of Governance seem to process personal data the most often, 
followed by the department of Law. Of course, it is possible that the other 
departments process more personal data than the respondents indicated. 

```{r leg-personaldataplot}
# Graph
leg_datatypegraphs <- leg_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_REBO_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 3,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across LEG departments") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

leg_datatypegraphs
```

### Geo

As with the university-wide data, researchers from the faculty of Geosciences
indicated that they also most frequently used textual and tabular data. However,
in contrast to the university-wide data, they are followed by geographical data
and code/theoretical models. This is to be expected considering the faculty. In 
terms of *personal* data, however, the same top-3 types are present here as in the 
entire university (demographic information, contact information, 
direct identifiers). 

```{r geo-datatypes, results = "asis"}
# Table of data types
geo_datatypetable <- dppsurvey %>% 
  filter(!is.na(Faculty_1)) %>%
  pivot_longer(cols = grep("^Datatype_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Datatype") %>%
  select(-name) %>%
  group_by(Datatype) %>%
  summarise(Count = length(Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Datatype = factor(Datatype, levels=Datatype)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)")) %>%
  select(-Count, -Percentage)

# Table of personal data types
geo_personaldatatypes <- dppsurvey %>%
  filter(!is.na(Faculty_1)) %>%
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  group_by(Personal_Datatype)

geo_tablepersdata <- geo_personaldatatypes %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype), # Update factor levels
         Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) %>% # reverse order
  mutate(Frequency = paste0(Count, 
                            " (",
                            Percentage,
                            "%)"),
  ) %>%
  select(-Count, -Percentage)

kables(list(
  kbl(geo_datatypetable,
      col.names = gsub("_", 
                       " ", 
                       names(geo_datatypetable)), 
      align = "l",
      caption = "<b>Types of data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F, 
                  html_font = "Verdana",
                  font_size = 12,
                  position = "float_left"),
  kbl(geo_tablepersdata, 
      col.names = gsub("_", 
                       " ", 
                       names(geo_tablepersdata)), 
      align = "l",
      caption = "<b>Types of <i>personal</i> data used</b>",
      valign = 't') %>%
    kable_classic(latex_options = "hover",
                  full_width = F,
                  html_font = "Verdana",
                  font_size = 12,
                  position = "left")))
```

When comparing departments, it becomes immediately clear that most personal
data seems to be processed in the department of Sustainable Development, and
some also in the Human Geography and Spatial Planning department. This makes 
sense, as the research performed at the third department (Earth sciences) usually
does not focus on human behaviour and thus does not involve much personal data, 
if any. 

```{r geo-personaldataplot}
# Graph
geo_datatypegraphs <- geo_personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_Geo_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 3,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across the Geosciences faculty") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

geo_datatypegraphs
```

## {-}

## Current practices in handling personal data

The first part of the survey addressed the researchers' current practices in 
handling personal data in their research.

### Organisational and technical measures {.tabset}
#### University-wide results
With respect to organisational and technical measures used to handle personal 
data, most respondents indicated that they pseudonymise/anonymise their data. 
As this question is self-reported, we cannot assess whether the researchers' data 
was actually sufficiently pseudonymised/anonymised in line with the GDPR. 
Secondly, many researchers seemed to implement access control and encryption, 
and complete a Data Management Plan (DMP) during their project(s). This is to
be expected, as most funders nowadays require a DMP and many DMP templates 
explicitly address topics like pseudonymisation, access control and encryption. 
On the other hand, GDPR-specific assessments such as Data Protection Impact 
Assessments (DPIAs) or privacy reviews were least used - as it stands, these 
assessments are still only carried out on a case-by-case basis.
 
```{r measuresused}
measures_plot <- 
  dppsurvey %>% 
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  pivot_longer(cols = grep("^Orgmeasures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Measures") %>%
  select(-name) %>%
  group_by(Measures, Faculty) %>%
  summarise(Count = length(Measures)) %>%
  arrange(Count) %>% # Order by count
  #mutate(Measures = factor(Measures, levels=Measures)) %>% # Update factor levels
  ggplot(aes(reorder(Measures, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Faculty)) +
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  labs(x = "", title = "Protective and planning measures",
       caption = "Organisational and technical measures used in handling personal data") +
  coord_flip() +
  theme_classic() +
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  geom_label(aes(label = after_stat(y), group = Measures),
             stat = 'summary',
             fun = sum,
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             label.size = NA, #1,
             nudge_y = 2,
             label.padding = unit(0.35, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  geom_text(aes(label = Count), # counts per measure per faculty
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))

measures_plot
```


#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Storage media used {.tabset}
#### University-wide results
When working with personal data, it is important to choose a sufficiently 
secure storage medium. Luckily, most respondents indicated to rely on storage 
solutions that are provided or recommended by UU - and in most cases are indeed
safe for storing personal data. Nonetheless, some respondents indicated to use 
non-UU solutions, including cloud solutions that UU advises against using. 
Several respondents also indicated that they use other storage solutions, such 
as those of external institutions (e.g., University Medical Center, Trimbos 
Institute, Central Bureau of Statistics) and repositories (e.g., DANS EASY, CLARIAH).

```{r storagemediaplot}
storage_used_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Storage_medium_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Storage") %>%
  select(-name) %>%
  group_by(Storage) %>%
  summarise(Count = length(Storage)) %>%
  arrange(Count) %>% # Order by count
  mutate(Storage = factor(Storage, levels=Storage)) %>% # Update factor levels
  ggplot(aes(x = Storage, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Storage media used",
       caption = "Storage Media Used for Storing Personal Data") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

  print(storage_used_plot)
```

#### Faculty-specific results
Here are some specific results

### {-}

### Consent forms: usage and content {.tabset}
#### University-wide results
Consent is currently the legal basis most often used in researcher performed at 
UU. Luckily, only a small number of respondents indicated that consent forms 
are not applicable to their research or they don't know what they are. The consent
forms most often used seem to be approved by privacy officers or ethical 
committees. Some respondents however indicated to independently create and use 
consent forms. A handful of respondents indicated to use templates from external 
institutions, or no templates at all (e.g., because they use verbal consent or 
data from another party).

The mere fact that researchers make use of a consent form and information letter 
of course does not mean these forms are of sufficient quality. Therefore, we 
asked respondents what they typically include in their information to 
participants. Most respondents indicated to provide at least the basic 
information on the purposes for data collection and which data will be be used. 
Information on the data subjects' rights and safeguards are often included as 
well. 

Information and consent on data sharing and reuse is apparently not included as 
often. This is a point of concern, because not obtaining consent for sharing 
personal data for reuse may mean that data collected for a specific research
project may not be reused again, reducing the scientific value of the data.

Finally, contact details of the Data Protection Officer (DPO) are not always 
included either, whereas these always have to be provided when working with
personal data (<a href = "https://gdpr-info.eu/art-13-gdpr/" target = "_blank">art. 13</a>)

```{r consentforms}
consent_usage_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consentforms_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Consent_Forms") %>%
  select(-name) %>%
  group_by(Consent_Forms) %>%
  summarise(Count = length(Consent_Forms)) %>%
  arrange(Count) %>% # Order by count
  mutate(Consent_Forms = factor(Consent_Forms, levels=Consent_Forms)) %>% # Update factor levels
  ggplot(aes(x = Consent_Forms, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Consent Forms",
       caption = "Usage of consent forms") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

consent_content_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Consent_content_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Consent_Content") %>%
  select(-name) %>%
  group_by(Consent_Content) %>%
  summarise(Count = length(Consent_Content)) %>%
  arrange(Count) %>% # Order by count
  mutate(Consent_Content = factor(Consent_Content, levels=Consent_Content)) %>% # Update factor levels
  ggplot(aes(x = Consent_Content, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Content of forms",
       caption = "Typical content of an informed consent form") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

#print(consent_content_plot)
grid.arrange(consent_usage_plot, consent_content_plot, nrow = 1)
```

#### Faculty-specific results
Here are some specific results

### {-}

### Data Protection Impact Assessment (DPIA) {.tabset}
#### University-wide results
A Data Protection Impact Assessment (DPIA) is a legal instrument to assess the
risks involved for data subjects, and helps determine the necessary safeguards to
reduce those risks to an acceptable level. Despite it being an important legal
instrument, most respondents indicated to never have carried one out, or heard of it,
for that matter. A minority of the sample had heard of it, or has completed one. 
Currrently, the desired scenario is that researchers get help from a privacy 
officer when performing a DPIA, and the results seem to indicate that this is
indeed the case in the majority of cases.

```{r dpia}
DPIA_experience_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_experience_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "DPIA_experience") %>%
  select(-name) %>%
  group_by(DPIA_experience) %>%
  summarise(Count = length(DPIA_experience)) %>%
  arrange(Count) %>% # Order by count
  mutate(DPIA_experience = factor(DPIA_experience, levels=DPIA_experience)) %>% # Update factor levels
  ggplot(aes(x = DPIA_experience, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Experience with DPIAs",
       caption = "Familiarity and experience with DPIAs") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

DPIA_help_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^DPIA_Help_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "DPIA_help") %>%
  select(-name) %>%
  group_by(DPIA_help) %>%
  summarise(Count = length(DPIA_help)) %>%
  arrange(Count) %>% # Order by count
  mutate(DPIA_help = factor(DPIA_help, levels=DPIA_help)) %>% # Update factor levels
  ggplot(aes(x = DPIA_help, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Help received with DPIAs",
       caption = "Help received with DPIAs") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(DPIA_experience_plot, DPIA_help_plot, nrow = 1)
```

#### Faculty-specific results
Blabla

### {-}

### Data sharing practices {.tabset}
#### University-wide results
To investigate how often data are being shared and under which circumstances, we
asked respondents if and with which parties they typically share their research
data, and which measures they usually take to do so securely. While there were 
some respondents who indicated not to share their data at all, most respondents
seem to only share research data within the organisation, and otherwise within the
European Economic Area (EEA). This is relatively good news, as such transfers 
usually require little, if any, additional safeguards. Responses in the "Other" 
category, however, suggest that a lot of data actually are shared, for example 
with co-authors at another institution, students, or in "pseudonymised form".

Concerning the measures used before sharing data, most repondents indicated to
pseudonymise their data, though we cannot assess the quality of such 
pseudonymisation. Researchers also seemed to use approved tools, agreements, 
and providing access without data transfer. Only a few respondents indicated that 
they involved a data expert while transferring data, and the use of Standard 
Contractual Clauses appears limited.

```{r datasharing}
external_sharing_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Share_outside_UU_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "external_sharing") %>%
  select(-name) %>%
  group_by(external_sharing) %>%
  summarise(Count = length(external_sharing)) %>%
  arrange(Count) %>% # Order by count
  mutate(external_sharing = factor(external_sharing, levels=external_sharing)) %>% # Update factor levels
  ggplot(aes(x = external_sharing, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "External sharing",
       caption = "Data Sharing with External Parties") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

sharing_measures_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Share_measures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "sharing_measures") %>%
  select(-name) %>%
  group_by(sharing_measures) %>%
  summarise(Count = length(sharing_measures)) %>%
  arrange(Count) %>% # Order by count
  mutate(sharing_measures = factor(sharing_measures, levels=sharing_measures)) %>% # Update factor levels
  ggplot(aes(x = sharing_measures, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Sharing measures",
       caption = "Protection measures taken while sharing data") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(external_sharing_plot, sharing_measures_plot, nrow = 1)

```

#### Faculty-specific
Blabla

### {-}

### Data publishing {.tabset}
#### University-wide

```{r datapublishing}
data_publishing_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Data_publication_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "data_publishing") %>%
  select(-name) %>%
  group_by(data_publishing) %>%
  summarise(Count = length(data_publishing)) %>%
  arrange(Count) %>% # Order by count
  mutate(data_publishing = factor(data_publishing, levels=data_publishing)) %>% # Update factor levels
  ggplot(aes(x = data_publishing, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Data Publishing",
       caption = "Data Publishing: Frequency and Content") +
  aes(stringr::str_wrap(data_publishing, 15), Count) + xlab(NULL) +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

print(data_publishing_plot)
```

Open science and privacy are often seen as conflicting, as sharing personal data 
cannot be done just like that, but requires at the very least a valid legal 
basis and additional safeguards. Therefore, in our experience, to date not many 
datasets that contain personal data are shared for reuse purposes, and the survey
respondents seemed to confirm this experience, as the majority of the respondents
indicated not to publish their data, or only in anonymised form (again, we cannot
be certain whether the data were indeed entirely anonymised).

The primary reason for not publishing data appears to be that researchers were 
still working on the data, they don't want to/need to publish their data, or 
they cannot anonymise the data. Other reasons given by the respondents included 
that publishing data is too much effort, publication is undesirable, or just not 
considered.

```{r publicationformat}
publication_format_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Publication_format_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "publication_format") %>%
  select(-name) %>%
  group_by(publication_format) %>%
  summarise(Count = length(publication_format)) %>%
  arrange(Count) %>% # Order by count
  mutate(publication_format = factor(publication_format, levels=publication_format)) %>% # Update factor levels
  ggplot(aes(x = publication_format, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Published Data Format",
       caption = "Format of Published Data") +
  aes(stringr::str_wrap(publication_format, 15), Count) + xlab(NULL) +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

reasons_nopub_plot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Reason_NoPub_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "reasons_nopub") %>%
  select(-name) %>%
  group_by(reasons_nopub) %>%
  summarise(Count = length(reasons_nopub)) %>%
  arrange(Count) %>% # Order by count
  mutate(reasons_nopub = factor(reasons_nopub, levels=reasons_nopub)) %>% # Update factor levels
  ggplot(aes(x = reasons_nopub, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", 
       title = "Why Not Published?",
       caption = "Reasons for Not Publishing Data") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(publication_format_plot, reasons_nopub_plot, nrow = 1)
```

#### Faculty
asdfasdf

### {-}

## Existing support channels
The second part of the survey concerned the visibility and use of existing 
support channels. 

### Faculty privacy officer
When asked whether respondents know who their faculty
privacy officer is, a little over half of the respondents indicated that they do 
not (Yes: `r sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%), No: 
`r sum(dppsurvey$Know_PO=="No", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%)).

When comparing faculties (see below), it is striking that the faculties where the
most personal data seems to be processed, the majority of respondents is not 
aware of their faculty privacy officer. This suggests either that researchers 
have simply never required help from their privacy officer, or that the faculty 
privacy officers could increase their visibility within their faculties.

```{r knowpoplot}
knowpoplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Faculty_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty, Know_PO) %>%
  summarise(Count = length(Know_PO)) %>%
  arrange(Count) %>% # Order by count
  filter(!is.na(Know_PO)) %>%
  mutate(Know_PO = factor(Know_PO, levels=Know_PO)) %>% # Update factor levels
  ggplot(aes(x = Know_PO, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Faculty, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", 
       title = "Awareness of the privacy officer",
       caption = "Awareness of the privacy officer across UU faculties") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2,
             label.size = NA, #1,
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

knowpoplot

```

### Looking for help {.tabset}
#### University-wide
When asked whether respondents have ever looked for support in the form of 
information, tools, or in-person support, an overwheling majority indicated that 
they have, as can be seen in the graph below. Most respondents that looked for support
indicated, however, that they haven't always found the support they were looking for.
Together with the results from the previous question, this suggests that the 
visibility of the current support channels could be improved.

```{r existingsupport}
# Ever looked for help with privacy?
lookedforhelpplot <- dppsurvey %>% 
  pivot_longer(cols = Lookedforhelp, 
               values_drop_na = TRUE,
               values_to = "searchhelp") %>%
  select(-name) %>%
  group_by(searchhelp) %>%
  summarise(Count = length(searchhelp)) %>%
  arrange(Count) %>% # Order by count
  mutate(searchhelp = factor(searchhelp, levels = searchhelp)) %>% # Update factor levels
  ggplot(aes(x = searchhelp, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Has looked for help",
       caption = "Have you ever looked for UU-specific information, \n in-person support or tools on handling personal data?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

# Found what you're looking for?
foundhelpplot <- dppsurvey %>% 
  pivot_longer(cols = Foundhelp, 
               values_drop_na = TRUE,
               values_to = "foundhelp") %>%
  select(-name) %>%
  group_by(foundhelp) %>%
  summarise(Count = length(foundhelp)) %>%
  arrange(Count) %>% # Order by count
  mutate(foundhelp = factor(foundhelp, levels = c("(Almost) always", 
                                                  "Often",
                                                  "Sometimes",
                                                  "(Almost) never"))) %>% # Update factor levels
  ggplot(aes(x = foundhelp, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Has found what they were looking for",
       caption = "Did you find what you were looking for \n(e.g., information, support, tools)?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

grid.arrange(lookedforhelpplot, foundhelpplot, nrow = 1)
```

#### Faculty-specific
Hello world

### {-}

### Channels used to find support {.tabset}
#### University-wide
The graph below indicates which channels respondents used most 
when looking for information about handling personal data. Note that this was 
a multiple choice question, and so the numbers do no add up to the total number
of survey respondents. As shown below, all sources mentioned were to some extent
consulted by the respondents. However, there is large differentiation between 
these sources. The university website and intranet are the most visited online 
resources for information about handling personal data. Notably, colleagues 
appear to play an important role as well in informing researchers
about how to handle personal data. This suggests that the better informed 
researchers are, the more positive the effect is on their colleagues as well. 
Moreover, this also suggests that in-person support may be a more effective way 
of increasing awareness of privacy-related practices than more "distant" 
information sources. 

```{r sourcesused}
# Sources used to find information and help
sourcesusedplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Sourcesused_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "sourcesused") %>%
  select(-name) %>%
  group_by(sourcesused) %>%
  summarise(Count = length(sourcesused)) %>%
  arrange(Count) %>% # Order by count
  mutate(sourcesused = factor(sourcesused, levels = sourcesused)) %>% # Update factor levels
  ggplot(aes(x = sourcesused, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Sources used to find help") +
  coord_flip() +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

sourcesusedplot
```

#### Faculty

### {-}

## Challenges and needs (survey)

```{r obstacle}
obstacleplot <- dppsurvey %>% 
  pivot_longer(cols = Obstacle, 
               values_drop_na = TRUE,
               values_to = "obstacle") %>%
  select(-name) %>%
  group_by(obstacle) %>%
  summarise(Count = length(obstacle)) %>%
  arrange(Count) %>% # Order by count
  mutate(obstacle = factor(obstacle, levels = c("(Almost) always",
                                                "Sometimes",
                                                "(Almost) never",
                                                "Not sure"))) %>% # Update factor levels
  ggplot(aes(x = obstacle, y = Count)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", title = "Is privacy an obstacle?",
       caption = "Do you see dealing with personal data as an obstacle in open science and research data management?") +
  theme_classic() + 
  geom_label(aes(label = Count), position=position_dodge(width=1.2),
             color = "black",
             fill = "#FFCD00")

obstacleplot
```

As can be seen above, most researchers experience privacy to be an obstacle for 
open science and research data management in some way. It is therefore important 
to aim for support in this area. What this support should look like, however, 
differs a bit. As can be seen below, accessible information and visible support
channels seem to be the most wanted improvements in the current support, closely
followed by UU-wide policy on the topic, and privacy-related walk-in hours. 

```{r bettersupport}
bettersupporttable <- dppsurvey %>% 
  pivot_longer(cols = grep("^Better_support_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "bettersupport") %>%
  select(-name) %>%
  group_by(bettersupport) %>%
  summarise(Count = length(bettersupport)) %>%
  arrange(Count) %>% # Order by count
  mutate(bettersupport = factor(bettersupport, levels=bettersupport)) %>% # Update factor levels
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  map_df(rev) #%>% # reverse order
  #print()

 kable(bettersupporttable, 
        col.names = gsub("_", " ", names(bettersupporttable)), 
        align = "l",
        caption = "How can we improve personal data-related services?")
```

## Challenges and needs (open questions, one-on-one meetings)

As mentioned in the [Methods section](#analysis), the responses on the following 
open questions, and the notes taken during the one-on-one meetings with 
researchers were coded to allow for easier analysis. From the survey, the 
following questions were coded:

- "Which challenges concerning the handling of personal data do you run into 
most often?"
- "What specific **information** or **tools** about handling personal data are 
you missing from existing sources?"
- "What can we do better to support you in handling personal data in research?" 
(responses to the "Other" option)

The codes that were assigned in both the survey and the meeting notes can be 
seen in the word cloud below: the larger the font, the more often the code 
was assigned. The meaning of each code is explained in the file 
<a href = "../data/processed/codes-freq_survey_meetings.csv" target = "_blank">"codes-freq_survey_meetings.csv"</a> 
and below. Please note that there is overlap in respondents between the open 
questions in the survey and the meeting notes, and therefore some codes may have 
been applied twice for the same researcher. 

```{r opentextcloud}
# Open most recent open text responses and coded meeting notes
opentextfile <- "../data/coded/dppsurvey_opentextresponses_coded.xlsx"
opentext <- read_excel(opentextfile, sheet = "opentextresponses")
opencodes <- read_excel(opentextfile, sheet = "codes")
opentools <- read_excel(opentextfile, sheet = "tools")

interviewcodes <- read_excel(interviewfile, sheet = "codes")
interviewtools <- read_excel(interviewfile, sheet = "tools")

# Merge open text responses and meeting notes codes (used the same codes)
totalcodes <- merge(opencodes, 
                    interviewcodes, 
                    by = c("Code", "Meaning"),
                    suffixes = c("_survey","_interview"))

# Count total times mentioned across survey and interviews
totalcodes <- totalcodes %>%
  mutate(Times_mentioned_total = `Times mentioned_survey`+ `Times mentioned_interview`) %>%
  arrange(desc(Times_mentioned_total))

# Prepare wordcloud
totalcodes_wc <- totalcodes %>%
  select(Code, Times_mentioned_total) %>%
  mutate(word = Code,
         freq = Times_mentioned_total,
         .keep = "none")

library(wordcloud2)
wordcloudcodes <- wordcloud2(data = totalcodes_wc, 
                             color=rep_len(UU_pallette, nrow(totalcodes_wc)),
                             fontFamily = "Open Sans",
                             minRotation = 0, 
                             maxRotation = 0,
                             rotateRatio = 1)
wordcloudcodes

# Save coded dataset to processed folder
write.csv(totalcodes, "../data/processed/codes-freq_survey_meetings.csv")

# For in-text reference, count total times certain tools are mentioned
totaltools <- merge(opentools, 
                    interviewtools, 
                    by = c("Name tool", "Explanation"),
                    suffixes = c("_survey","_interview"))

# Count total times mentioned across survey and interviews
totaltools <- totaltools %>%
  mutate(Times_mentioned_total = `Times mentioned_survey`+ `Times mentioned_interview`) %>%
  mutate(Name_tool = `Name tool`) %>%
  select(-`Name tool`) %>%
  arrange(desc(Times_mentioned_total))
```

**Below we go into the most frequently mentioned challenges/needs expressed
by researchers in the survey (open questions) and one-one-one meetings:**

### Visibility and findability
`r totalcodes$Meaning[totalcodes$Code == "visibility"]` (mentioned 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "visibility"]` times). 
Some respondents indicated not to know where to go for help with privacy-related
matters, others mentioned that there should be *more* support personnel available
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "availability"]`). 
One researcher said to prefer having one place to go to for all
data-related questions. As a matter of fact, many faculties have such a 
"one-stop-shop" in place, but responses such as these indicate that their 
visibility may be improved.

Closely related to visibility is findability. Many researchers point out that 
available information is difficult to find, or it is confusing which source 
should be followed. As an example, one researcher from the faculty of Humanities
was determined that they found some useful information at the RDM Support website.
However, upon closer inspection, it appeared that this information arose from
another data management-related website with the UU logo on it. 

<blockquote>"UU biedt ontzettend veel aan, maar je moet veel websites bezoeken 
om alles te vinden [...] Een duidelijk overzicht voor welke informatie je
waar moet zijn zou fijn zijn."</blockquote>

### Specific tools
Some (`r totalcodes$Times_mentioned_total[totalcodes$Code == "tools"]`) 
researchers expressed a need for a concrete (improvement of a) tool, such 
as for:

- `r totaltools$Explanation[totaltools$Name_tool == "transcription"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "transcription"]` 
times)
- `r totaltools$Explanation[totaltools$Name_tool == "computing-critical"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "computing-critical"]` 
times)
- `r totaltools$Explanation[totaltools$Name_tool == "pseudonymising-video"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "pseudonymising-video"]` 
times)

### Hands-on support
Many (`r totalcodes$Times_mentioned_total[totalcodes$Code == "hands-on"]`) 
researchers indicated that support staff could sometimes provide support 
in a more hands-on fashion, rather than abstract advice and telling researchers 
how *not* to do things. Some researchers added that privacy professionals sometimes
have a tendency to cling to the letter of the law, leading to significant
delays in their project, instead of looking at how to concretely solve existing 
issues in practice (mentioned `r totalcodes$Times_mentioned_total[totalcodes$Code == "strict"]` 
times): 

<blockquote>"Soms zijn we door deze regels heiliger dan de paus."</blockquote>
<blockquote>"Actual getting-your-hands-dirty support: not the kind that tells 
you what to do, but also the kind that helps you by doing."</blockquote>

### Less bureaucracy
Processes were often experienced as time-inefficient, and sometimes longer 
and more bureaucratic than necessary (mentioned by 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "quicker"]` researchers).
For example, the DPIA process was mentioned explicitly 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "dpia-process"]` times),
as well as having to fill out too many forms with overlapping content (e.g.,
Privacy Scan, Data Management Plan, DPIA). Some 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "divide-responsibilities"]`) 
researchers argued that (part of) this burden should be relieved or 
carried by support staff:

<blockquote>"Minder acties die gericht zijn op inhoudelijk trainen van WP en 
meer uit handen nemen van deze groep."</blockquote>
<blockquote>"Sharing data costs a lot of time and is inefficient when you do 
not do it often"</blockquote>

### Unclear processes and guidelines
Many (`r totalcodes$Times_mentioned_total[totalcodes$Code == "unclear-process"]`) 
researchers complained that it was unclear what was expected of them when 
they processed personal data in their research, or that they would like to have more,
or more practical guidelines (`r totalcodes$Times_mentioned_total[totalcodes$Code == "guidelines"]`)
on this topic, for example on:

- What steps do researchers need to take? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "assessment-steps"]`)
- Who should researcher ask for help in which situation?
- `r totalcodes$Meaning[totalcodes$Code=="responsibility"]` (`r totalcodes$Times_mentioned_total[totalcodes$Code == "responsibility"]`)
- `r totalcodes$Meaning[totalcodes$Code=="changing-rules"]` (`r totalcodes$Times_mentioned_total[totalcodes$Code == "changing-rules"]`)
- Different processes for different types of research (e.g., student projects vs.
large longitudinal projects, `r totalcodes$Times_mentioned_total[totalcodes$Code == "different-processes"]`)

### Information and education
A large part of the respondents expressed a need for more (clear) information 
and education with respect to handling personal data in research. In general, 
researchers indicated that the information offered to them should be 
more clear (`r totalcodes$Times_mentioned_total[totalcodes$Code == "unclear-info"]`), 
simpler `r totalcodes$Times_mentioned_total[totalcodes$Code == "info-simple"]`, 
consistent across resources (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-consistent"]`), 
possibly in the form of templates (`r totalcodes$Times_mentioned_total[totalcodes$Code == "templates"]`). 
Luckily, some researchers indicated to already be happy with existing materials (`r totalcodes$Times_mentioned_total[totalcodes$Code == "positive-existing-materials"]`).

#### Information for specific research or data types
Many researchers indicated to have a need for more tailored information for specific
types of data or research 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-specific"]`), 
for example for etnographic data 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-etnography"]`),
historical data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-history"]`), 
or video data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-video"]`).

<blockquote>"De informatie is gewoon veel te generiek, er zouden templates 
moeten zijn per type onderzoek."</blockquote>
<blockquote>"Veel templates en uitlegmodellen spreken over data, data packages 
en metadata, maar die woorden zijn niet ingebed in historisch onderzoek. Er 
ontstaat al snel verwarring over wat historici nu precies moeten met 
archiefmateriaal in het licht van privacy."</blockquote>

#### Frequently asked questions
A selection of researchers used the space in the open questions and/or the meetings 
to ask knowledge-related questions. These can be used in future support efforts,
such as the 
<a href="https://utrechtuniversity.github.io/dataprivacyhandbook" target="_blank">Data 
Privacy Handbook</a>, future courses, or other future information campaigns to 
better address researchers' needs. Below are examples of the most commonly 
asked questions:

- When are data still personal (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-personal-data"]`), 
how to anonymise data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "anonymise-how"]`),
and when are data anonymised sufficiently (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-anonymous"]`)?
- What are the privacy-related requirements for students? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "students"]`)
- How to store different types of personal data? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "datastorage"]`)
- Data sharing: what data can be shared and with whom? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "datasharing"]`)
- How to find a balance between informing data subjects too little vs. providing
too much privacy-related information that will scare them off or hurt their 
trust in research? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-informing-toomuch"]`)
- How to balance open science and privacy? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "open-vs-privacy"]`).
Related: when is reusing data fir different purposes allowed? (e.g., education 
data, data collected by students, `r totalcodes$Times_mentioned_total[totalcodes$Code == "info-reuse"][1]`)
- How to collaborate with multiple institutions? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "collaboration"]`)

#### Education and onboarding
The need for more educational resources was also recognised by a selection of the
respondents. Concretely, researchers could be educated more in the following ways:

- Privacy and/or research data management as part of the master or PhD 
curriculum (`r totalcodes$Times_mentioned_total[totalcodes$Code == "education"]`)
- Privacy as part of the onboarding procedure of new employees (`r totalcodes$Times_mentioned_total[totalcodes$Code == "onboarding"]`)
- Mandatory privacy training for supervisors, principal investigators, professors, 
and/or teachers (`r totalcodes$Times_mentioned_total[totalcodes$Code == "education-supervisors"]`)
- A course on how to handle personal data in research (`r totalcodes$Times_mentioned_total[totalcodes$Code == "course"]`)

<blockquote>"There is no one who tells at the start of your PhD how you should handle your data. [...] I think new PhD students should get a basic course on data management and privacy."</blockquote>

### ... or simply no issues
Notably, there were also researchers who indicated **not** to have run into issues 
(yet), or to have received sufficient and useful help (`r totalcodes$Times_mentioned_total[totalcodes$Code == "positive-existing-support"]`). For example:

<blockquote>"The data manager and privacy officer of the faculty of humanities help a lot. This support is essential!"</blockquote>
<blockquote>"Tot nu toe heb ik niet veel problemen gehad. De institutional review board van onze afdeling kijkt altijd kritisch naar de onderzoeksvoorstellen, ook met name op omgaan met persoonsgegevens."</blockquote>

# Conclusion
This report summarised the findings of the Data Privacy Survey and subsequent
one-on-one meetings with researchers. Our aim was to investigate the current
practices, challenges and needs of Utrecht  University (UU) researchers in the
handling of personal data. 

## Results summary
In short, we have found that:

- Personal data are processed in each faculty and across all academic positions
at Utrecht University, but mostly among the faculties of Science and Social and
Behavioural Sciences, and among early-career researchers.
- Most researchers deal with tabular, textual, code, and audiovisual data, which
contain mainly demographics, contact information, and direct identifiers.
- Most researchers seem to be at least familiar with concepts like anonymisation 
and pseudonymisation, access control, and UU-approved tools for handling personal 
data. However, the knowledge level differs quite a bit. For example:
  - some researchers seem to store data on locations that officially are not allowed.
  - information to study participants is sometimes lacking all the required information, 
  such as contact information from the Data Protection Officer, and information and
  consent for sharing data for reuse purposes.
  - about half of the sample is not familiar with the concept of a Data 
  Protection Impact Assessment.
  - data are often shared, but not all the right measures are always taken to 
  do so securely. 
- Privacy is sometimes found to be an obstable for open science, also confirmed by 
a relatively low amount of published datasets.
- Not all help is found equally well. For example, about half of the respondents
indicated not to be aware of their faculty privacy officer, and many researchers
resort to the UU website, intranet, and colleagues for help.
- Researchers may benefit from greater visibility and findability of support, 
specific tools. more and more hands-on support, more clear and less bureaucratic
processes and guidelines, and specific information and education.

Together, these findings suggest that there is still a lot to be gained
regarding awareness and secure practices in scientific research with personal data.

## Recommendations
### Increased visibility and findability
The findings show that the visibility of existing support for handling personal 
data can be increased at UU. This includes visibility of privacy officers, 
legal affairs and Research Data Management Support, as well as faculty- or 
department-level support. 

As shown by the fact that many respondents consult with colleagues for 
information and support, personal contact with supporters will likely be more 
effective at increasing awareness and uptake of secure privacy practices than a 
passive website that may not always answer all possible questions.

Some suggestions for increasing visibility and findability:

- Make sure there is (sufficient) support available to help researchers in time.
If there is no support available, the odds are that researchers may not ask for 
help at all and thus forget important privacy elements in the design of their 
research project.
- Make sure support staff are easy to find. For example, give them a clear location
on the university website and/or on the intranet. Also indicate who is available
for what type of support and for which target audience (e.g., "all UU 
researchers", "employees from the Faculty of Humanities", etc.)
- Make sure information on handling personal data in research is easy to find.
The UU website and intranet seem the most suitable places, as these are the most
visited to look for information according to the survey. 
- Make researchers actively aware of existing support channels. For example,
first-line support can visit important faculty or department events, offer to
give a workshop, and keep short lines of contact with department heads.

### More hands-on support
In line with the [findings described above](#hands-on-support), more hands-on 
support is desirable. This can for example be reached by:

- More collaboration between different types of data support. For example, at UU
there is a clear distinction between privacy officers (employed by faculties),
data managers (employed by the University Library and some faculties), information
security officers (Information Technology Services) and research engineers 
(Information Technology Services). However, in practice there is often overlap 
between either or all of these positions. Regular collaboration between these 
parties can:
  - help privacy and security experts understand the perspective of the 
  researcher and possibly how to apply their advice in practice,
  - help data managers and research engineers understand the GDPR requirements 
  and translate those to practice, and
  - enable reuse of solutions already applied in earlier cases that either 
  party may not be aware of.
- Together with researchers, developing suitable practical solutions (tools, 
protocols, etc.) for specific scenarios and/or types of research, and sharing 
them with the world for reuse (if possible).
- Increasing contact between support staff and researchers to understand the 
issues they are dealing with that require a solution.

### Decreased administrative burden

In line with the [findings described above](#less-bureaucracy), it is desirable
to attempt to decrease the administrative burden on researchers. We suggest that 
the bureaucracy issues can be partly attributed to:

1. Understaffing: it is currently difficult for data support staff to review 
*all* research projects involving personal data at UU. This can lead to 
projects with issues being noticed too late and researchers having 
to make changes to their project after the fact. This creates unnecessary 
administrative burden that could have been prevented had researchers been 
in contact with support staff from early on. 
2. Lack of unifying policy on steps to take when working with personal data: At
the moment each faculty has its own procedure, or has none at all. This can 
create confusion among both researchers and support staff on which steps to take
when, who to contact for help, or who is responsible for deciding on compliance.
3. Lack of integration between systems and staff: having to fill out the same
thing multiple times may be caused by the fact that different administrative
systems are used for different purposes. Integration of these systems may lead
to a relief of the administrative burden experienced by researchers. <br>For 
example, DMPonline is used at UU for registering Data Management Plans. However, 
DMP templates do not include sufficient questions about how personal data are 
handled, and thus a privacy scan may be additionally required. Privacy is 
sometimes part of an ethical application, and so this same information may be 
included there again as well. Finally, the processing register also requires 
information on (among others) the type and sensitivity of personal data, 
which is a third system where this information will be registered.

### Education and outreach

Many problems seem to arise from the fact that privacy has not been taken into
account from the design phase of a research project. This leads to research
support staff being contacted too late in the process, who then has to tell the
researcher to retrospectively make changes in their design in order to move 
forward. Such scenarios may be prevented by, for example:

- Including the basic of handling personal data and contact details of support
staff in onboarding processes when researchers are newly hired.
- Educating researchers on taking privacy into account in the designing phase
of their project. 
- Organising outreach events, such as the "Privacy and Security week" making
researchers aware of privacy-related measures (see further Visibility).
- Embedding privacy in existing materials and processes, such as in ethical
applications, in the education of research master and PhD students, etc.

### Specific information and tools
The specific [information](#frequenty-asked-questions) and 
[tools](specific-tools) described above can be addressed in current and future
support efforts. One of these is the 
<a href="https://utrechtuniversity.github.io/dataprivacyproject">Data Privacy Project</a>, 
from which this Data Privacy Survey forms a part. Other efforts may include the
updating of websites from RDM Support, faculties, or ethical committees, and the
updating of services offered by the IT department.

## Discussion

In order to interpret the results described in this report correctly, there 
are a few matters that need to be taken into account:

1. The respondents of the survey represented only a small selection of all of 
UU's Scientific Personnel. Therefore, the results presented here should not be 
overgeneralised to represent all of UU's researchers. It is likely that at least 
part of the respondents had a specific reason to do so. For example, they may 
have explicitly wanted to raise their voice because of some bad experiences 
they had. Although it is important that these concerns be heard, they should not 
be interpreted as to speak for all UU researchers.   
2. It is important to note that we, the writers of this report, cannot be 
considered fully objective in interpreting the survey results. The feedback 
received via this survey also involves our own services and possibly work 
performance. Although we have attempted to interpret the results of the survey 
without bias, it is possible that we have indeed introduced some, especially 
where the open questions and one-on-one meetings are concerned.   
3. The questions in this survey intentionally do not address at which data 
support personnel the feedback is targeted. This also means that our 
recommendations do not specify Who within the organisation should pick them up. 
Preferably, these recommendations will be picked up by a  collaboration between 
data support at different levels of the organisation (departmental, faculty, 
central) and, if needed, beyond, to address researchers' concerns and improve 
the support surrounding the handling of personal data in research.

# Technical information
This document was last created on: `r as.character(Sys.Date())`. It was 
created in the following environment:

```{r sessioninfo}
sessionInfo()
```
