---
title: "Data Privacy Survey"
subtitle: "Challenges and needs of privacy-related services for research at Utrecht University"
author:
- "Dorien Huijser, Research Data Management Support, Utrecht University"
- "Neha Moopen, Research Data Management Support, Utrecht University"
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
abstract: |
  In the second quarter of 2022, Utrecht University (UU) Research Data Management 
  Support (RDM Support) sent out a survey among all scientific personnel at UU: 
  the Data Privacy Survey. The aim of this survey was to investigate 1) How UU 
  researchers currently deal with personal data in research, 2) What challenges 
  they run into when handling personal data in research, and 3) How support at
  UU can improve their services concerning personal data in research. The 
  results show that most researchers know to take into account 
  privacy in their projects. However, there are vast differences in knowledge 
  on this topic, as well as confusion on topics such as data sharing, 
  anonymisation, and the administrative requirements surrounding privacy. Taken
  together, the survey shows that researchers could benefit from research support
  that is more visible in the organisation and more hands-on (i.e., advise that is
  immediately applicable instead of more general). Additionally, more awareness 
  of the most important privacy-related topics, and how to deal with them in 
  research seems desirable, both for newly employed and existing researchers.
  Finally, many researchers indicate to experience a high administrative burden 
  that is partly caused by privacy-related documentation before starting a 
  project. Finding a way to decrease this burden while still complying with 
  legal and policy regulations will likely help to increase compliance and 
  help researchers focus on performing high-quality research.
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
    df_print: paged
    css: style.css
    includes:
      in_header: "../includes/header.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

```{r sourcescript, include=FALSE, cache=FALSE}
knitr::read_chunk('plot-data.R')
source("plot-data.R", local = knitr::knit_global())
```

# Introduction {#introduction}
Scientific research often includes some form of personal data. The amount of 
personal data that is processed in research is likely underestimated, as 
researchers may be unaware of what personal data are or whether they are being 
collected. With the implementation of the General Data Protection Regulation
(GDPR) in 2018, stricter legal requirements apply to handling personal data and 
its sharing and publication. In our own experience, the number and complexity of 
questions on handling personal data in scientific research at Utrecht University
(UU) is increasing. 

Our premise at <a href = "https://uu.nl/rdm" target = "_blank">Research Data 
Management Support</a> (RDM Support) is to assist researchers with any issues 
surrounding the management of their research data, including research data that 
contains personal data. To understand how we can best help researchers with 
their privacy-related questions and needs, we wanted to investigate:

1. To what extent are UU researchers aware of privacy legislation and practices?
2. What data privacy issues do UU researchers typically run into?
3. What support do UU researchers need to handle personal data?

To answer these questions, we set up an online survey and planned one-on-one 
meetings with a selection of UU researchers. This report describes our methods, 
results and the future steps that data support at UU can take to improve their 
services concerning privacy in research.

This survey was part of a larger project at UU, the 
<a href = "https://utrechtuniversity.github.io/dataprivacyproject/" target = "_blank">Data Privacy Project</a>^[The 
Data Privacy Project is funded by Utrecht University’s Research IT program and 
a Digital Competence Center grant from the Dutch Organization for Scientific 
Research (NWO). It is led by Utrecht University's 
<a href="https://uu.nl/rdm/" target = "_blank">Research Data Management Support</a>, 
in collaboration with the University Library, Information Technology Services, 
first- and second-line privacy officers, information security, and others.]. 
This project is a data support effort within UU that aims to provide 
actionable and FAIR (Findable, Accessible, Interoperable, Reusable) information 
and tools for researchers to handle personal data in their research. 

# Methods {#methods}
```{r dependencies}
```

```{r readdata}
```

```{r labelsfsw}
# The labels of the FSW Departments are too long to display properly in a graph.
# Therefore they are shortened in this block
dppsurvey <- dppsurvey %>%
  mutate(Dept_FSW_3 = recode(Dept_FSW_3, 
                             "Development & Education of Youth in Diverse Societies" = "Pedagogy in Diverse Societies"),
         Dept_FSW_4 = recode(Dept_FSW_4,
                             "Clinical Child & Family Studies" = "Clinical Child & Family"),
         Dept_FSW_5 = recode(Dept_FSW_5,
                             "Interdisciplinary Social Science" = "Interdisc. Social Sci"),
         Dept_FSW_12 = recode(Dept_FSW_12, 
                              "Social Health & Organisational Psychology" = "Social, Health & Org."))
```

```{r facultydatasets}
# Read in faculty datasets
dppsurvey_science <- dppsurvey %>% filter(!is.na(Faculty_5))
dppsurvey_fsw <- dppsurvey %>% filter(!is.na(Faculty_6))
dppsurvey_geo <- dppsurvey %>% filter(!is.na(Faculty_1))
dppsurvey_vet <- dppsurvey %>% filter(!is.na(Faculty_7))
dppsurvey_leg <- dppsurvey %>% filter(!is.na(Faculty_3))
dppsurvey_hum <- dppsurvey %>% filter(!is.na(Faculty_2))
```

```{r uucolors}
```

## Survey questions {#survey-questions}
The [Data Privacy Survey](#data-and-material-availability) was created by the project 
coordinators (DH, NM) with input from a wide variety of experts, consisting of 
privacy experts, data managers, data consultants, and IT staff (e.g., information 
security,  research engineering). The survey consisted of 19-23 questions 
(the exact number differing depending on given answers), which were separated 
into 5 sections:

1. **About you and your research**: 5 questions about the faculty, department, 
position, and type of (personal) data the researcher works with.
2. **Measures and documents**: 4-6 questions about researchers' familiarity and
use of privacy-related measures (e.g., processing register, Data Privacy Impact
Assessment, encryption, pseudonymisation, etc.), data storage, and informed 
consent forms.
3. **Data sharing, archiving and publication**: 3-4 questions about researchers' 
data sharing and publishing practices.
4. **Finding support**: 3-4 questions about the awareness of several data support
channels.
5. **Improving our services**: 4 questions about researchers' challenges, needs, 
and suggestions to improve UU's data privacy support.

The estimated 10 minute duration of the survey in advance proved to be 
relatively accurate: respondents spent a median of 
`r round(median(dppsurvey$Duration/60), 1)` minutes on the survey (although 
there was a very large variation; range: 
`r round(min(dppsurvey$Duration)/60, 1)` minutes - 
`r round(max(dppsurvey$Duration)/3600/24,1)` days).

## Procedure online survey {#procedure-online-survey}
The survey was created in Qualtrics and distributed from March 21st, 2022 
onwards via several communication channels to reach as many UU researchers as 
possible:

- An <a href = "https://utrechtuniversity.github.io/dataprivacyproject/20220321-data-privacy-survey.html"
target = "_blank">email</a> was sent to all academic staff at UU through 
central communication channels. The reasoning was that this would be the most
effective way to reach as many UU researchers as possible - taking for granted
that we would likely miss a small part of non-academic personnel also involved
in research in some way.
- A mention in several Faculty newsletters.
- Social media messages (e.g., on <a href = "https://twitter.com/RDMsupportUU/status/1517390979707412481"
target = "_blank">Twitter</a>).
- A <a href = "https://intranet.uu.nl/en/news/news-items/survey-better-support-for-handling-privacy-sensitive-data-in-research" target = "_blank">news item</a> on the UU intranet.
- Via data support colleagues, who were asked to point researchers they were
in contact with to the survey.

Researchers could decide voluntarily whether they wanted to participate in the
survey. All results reported here are from researchers who provided their active 
consent, worked at Utrecht University and indicated to work with some type of 
research data.

## Online survey respondents {#online-survey-respondents .tabset}
### UU-wide
The survey was filled out by `r dim(dppsurvey)[1]` UU researchers. As can be 
seen in the figure below, we received responses from each UU faculty, but the 
distribution was not equal: the faculties of Science and Social and Behavioural 
Sciences were overrepresented, whereas the responses of the Faculty of 
Geosciences and the Faculty of Medicine were rather low. This can be explained 
by the mere size of the faculties (e.g., the faculty of Science is UU's largest 
faculty), but also by the types of research performed there. For example, 
research performed at the faculty of Geosciences is largely involved with 
natural scientific data, rather than data from or relating to humans. The 
Faculty of Medicine, then, is located at the University Medical Center Utrecht 
(UMCU) and is not of primary interest for the current survey, which was sent to 
UU researchers only. A small selection of respondents indicated to work at 
another part of the organisation, being mainly University College Utrecht and 
the University Corporate Office.

Most of the survey respondents were relatively early 
career researchers (PhDs, junior researchers and postdoctoral researchers). 
We suspect that this is because 1) there simply are more people in these positions
than there are in the others, and 2) early-career researchers may experience a 
greater need for support with respect to handling personal data.

```{r facultyplot}
# Plot of positions per faculty
position_per_faculty <- dppsurvey %>%
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Faculty, Position) %>%
  summarise(Count = length(Faculty)) %>%
  ggplot(aes(reorder(Faculty, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Position)) +
  coord_flip() +
  theme_classic() + 
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  labs(x = "", 
       title = "Faculty and Position representation",
       caption = "Survey respondents per faculty and position") +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  #scale_x_continuous(expand = expansion(add = 3)) + 
  geom_label(aes(label = after_stat(y), group = Faculty), # totals per faculty
             stat = 'summary', 
             fun = sum, 
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             nudge_y = 2,
             label.size = NA, #1,
             label.padding = unit(0.35, "lines"),
             #label.r = unit(0.25, "lines"),
             color = "black",
             fill = "#FFFFFF") + 
  geom_text(aes(label = Count), # counts per position per faculty
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))

position_per_faculty

```

### Science
There were `r dim(dppsurvey_science)[1]` respondents from the Science faculty 
in the online survey. It took them a median of 
`r round(median(dppsurvey_science$Duration/60), 1)` minutes to complete it
(ranging from `r round(min(dppsurvey_science$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_science$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Science faculty is visualised. 

```{r sciencedepartments}
plotdepartments(dppsurvey_science, "Science", 
                 title = "Departments and Positions Science", 
                 caption = "Survey respondents per department and position in the Science faculty")
```

Most respondents in the online survey seemed to come from the Information and
Computing Sciences (ICS) and Pharmaceutical Sciences departments, and were 
relatively early career researchers (PhD/junior/postdoc researcher, 
assistant/associate professor). This makes sense, as more personal data is 
likely processed in the fields of ICS and pharmaceuticals, compared to, for 
example, mathematics.

### FSBS
There were `r dim(dppsurvey_fsw)[1]` respondents from the Faculty of Social and
Behavioural Sciences (FSBS) in the online survey. It took them a median of 
`r round(median(dppsurvey_fsw$Duration/60), 1)` minutes to complete it (ranging 
from `r round(min(dppsurvey_fsw$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_fsw$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within FSBS is visualised. 

```{r fswdepartments}
plotdepartments(dppsurvey_fsw, "FSW", 
                 title = "Departments and Positions FSBS", 
                 caption = "Survey respondents per department and position at FSBS")
```

Most respondents in the online survey seemed to come from the Education department, 
although closely followed by Methodology and Statistics and Interdisciplinary 
Social Sciences. Additionally, respondents from the Faculty of Social and 
Behavioural Sciences were mostly either PhD/junior/post-doc researcher or 
Assistant/Associate professors. 

### Humanities
There were `r dim(dppsurvey_hum)[1]` respondents from the Faculty of Humanities
in the online survey. It took them a median of 
`r round(median(dppsurvey_hum$Duration/60), 1)` minutes to complete it (ranging 
from `r round(min(dppsurvey_hum$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_hum$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Faculty of Humanities is visualised. 

```{r humdepartments}
plotdepartments(dppsurvey_hum, "Humanities", 
                 title = "Departments and Positions Humanities", 
                 caption = "Survey respondents per department and position at the Faculty of Humanities")
```

The majority of Humanities researchers responding to the survey were employed 
at the department of Languages, Literature and Communicaiton. This makes sense, 
as likely a lot of personal data is processed there. Position-wise, there was 
quite some variability, although most respondents indicated to be either 
PhD/junior/postdoc researcher or Assistant/Associate Professor. 

### Veterinary Medicine
There were `r dim(dppsurvey_vet)[1]` respondents from the Faculty of Veterinary
Medicine in the online survey. It took them a median of 
`r round(median(dppsurvey_vet$Duration/60), 1)` minutes to complete it (ranging 
from `r round(min(dppsurvey_vet$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_vet$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Faculty of Veterinary Medicine 
is visualised. 

```{r vetdepartments}
plotdepartments(dppsurvey_vet, "DGK", 
                 title = "Departments and Positions Veterinary Medicine", 
                 caption = "Survey respondents per department and position at the Faculty of Veterinary Medicine")
```

Most respondents from the Veterinary Medicine faculty were employed at the 
department of Population Health Sciences. This makes sense as this is likely 
the department in which the largest amount of personal data is processed in 
research. Interestingly, the respondents consisted only of Assistant/Associate 
professors and PhD/junior/postdoc researchers, but also support staff and teachers.

### LEG
There were `r dim(dppsurvey_leg)[1]` respondents from the Faculty of Law, 
Economics, and Governance (LEG) in the online survey. It took them a median of 
`r round(median(dppsurvey_leg$Duration/60), 1)` minutes to complete it (ranging 
from `r round(min(dppsurvey_leg$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_leg$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Faculty of Law, Economics, and 
Governance (LEG) is visualised. 

```{r legdepartments}
plotdepartments(dppsurvey_leg, "REBO", 
                 title = "Departments and Positions LEG", 
                 caption = "Survey respondents per department and position at the Faculty of Law, Economics, and Governance")
```

Most LEG respondents came from the Governance department, closely followed by 
the department of Law. Respondents were mostly Assistant/Associate professors, 
with some PhD/junior/postdoc researchers and a few professors.

### Geo
There were `r dim(dppsurvey_hum)[1]` respondents from the Faculty of Geosciences
in the online survey. It took them a median of 
`r round(median(dppsurvey_geo$Duration/60), 1)` minutes to complete it (ranging 
from `r round(min(dppsurvey_geo$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey_geo$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Faculty of Geosciences is visualised. 

```{r geodepartments}
plotdepartments(dppsurvey_geo, "Geo", 
                 title = "Departments and Positions Geosciences", 
                 caption = "Survey respondents per department and position at the Faculty of Geosciences")
```

From the Geosciences faculty, most respondents were employed at the department 
of Sustainable Development, followed by Human Geography and Spatial Planning. 
Respondents were mostly PhD/junior/postdoc researchers and Assistant/Associate 
professors. Earth Sciences did not have many respondents, which makes sense as 
it is less likely that personal data are being processed there in the research 
context. 

### {-}

## One-on-one meetings {#one-on-one-meetings .tabset}
Besides the online survey, we organised one-on-one meetings with researchers, to 
hear about their personal experiences, challenges and needs concerning the handling 
of personal data in their research. Survey respondents could voluntarily leave 
their email address at the end of the survey to be contacted by us. These 
meetings were semi-structured and revolved around the following questions:

- What made you leave your email address in the Data Privacy Survey? Related:
What are your general experiences in handling personal data in research?
- Which difficulties do you run into when handling personal data?
- What support would you need to help you handle personal data in your research?
- Do you have a concrete need for support at the moment?

All of the one-on-one meetings were conducted online and took approximately 30
minutes. During the meetings, one of the project coordinators (DH) was always 
present, together with either the other project coordinator (NM) or the 
relevant faculty privacy officer. Before the privacy officer was invited to 
the meeting, the researcher's consent to do so was always obtained first.

```{r read-opentext}
```

```{r opentext-faculties}
# Filter interview and open text responses codes on Science faculty
interviews_science <- interviews %>% filter(grepl("Science(?!s)", Faculty, perl = TRUE))
interviews_fsw <- interviews %>% filter(grepl("Social and Beh*", Faculty, perl = TRUE))
interviews_leg <- interviews %>% filter(grepl("Law Economics", Faculty, perl = TRUE))
interviews_geo <- interviews %>% filter(grepl("Geosciences", Faculty, perl = TRUE))
interviews_vet <- interviews %>% filter(grepl("Veterinary", Faculty, perl = TRUE))
interviews_hum <- interviews %>% filter(grepl("Humanities", Faculty, perl = TRUE))

opentext_science <- opentext %>% filter(grepl("Science(?!s)", Faculty, perl = TRUE))
opentext_fsw <- opentext %>% filter(grepl("Social and Beh*", Faculty, perl = TRUE))
opentext_leg <- opentext %>% filter(grepl("Law Economics", Faculty, perl = TRUE))
opentext_geo <- opentext %>% filter(grepl("Geosciences", Faculty, perl = TRUE))
opentext_vet <- opentext %>% filter(grepl("Veterinary", Faculty, perl = TRUE))
opentext_hum <- opentext %>% filter(grepl("Humanities", Faculty, perl = TRUE))
```

### UU-wide
From the survey respondents, `r sum(dppsurvey$Email == "Yes", na.rm=TRUE)`
researchers left their email address to be contacted. Of those, 28 researchers 
indicated that they were willing to meet with us. Below, the division over 
faculties can be seen for all interviewees.

```{r emailaddressplot}
meetings_faculties <- data.frame(Faculty = str_trim(unlist(str_split(interviews$Faculty, 
                                                                     ",")), 
                                                    side = "both"))

meetings_faculties %>%
  group_by(Faculty) %>%
  summarise(Count = length(Faculty)) %>%
  arrange(Count) %>% # Order by count
  mutate(Faculty = factor(Faculty, levels=Faculty)) %>% # Update factor levels
  mutate(Percentage = round(Count/sum(Count)*100,0)) %>%
  arrange(Percentage) %>% # Order by %
  ggplot(aes(x = Faculty, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  labs(x = "", 
       title = "Faculty representation one-on-one meetings",
       caption = "Faculty representation of researchers spoken to in one-on-one meetings") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = paste0(Percentage, "%")), position=position_dodge(width=1.2),
             vjust = 0.5, 
             hjust = 'inward', #-0.2,
             label.size = NA, #1,
             label.padding = unit(0.35, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  theme(axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank())

```

### Science

From the survey respondents, `r sum(dppsurvey_science$Email == "Yes", na.rm=TRUE)` 
researchers from the Science faculty left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews_science)[1]` online meetings have been 
conducted, each with the project coordinators. Below, the division over 
positions can be seen for all interviewees from the Science faculty.

```{r positionsmeetings-science}
positionsmeetings(interviews_science,
                  title = "Positions one-on-one meetings Science",
                  caption = "Positions of Science researchers in one-on-one meetings")
```

### FSBS

From the survey respondents, `r sum(dppsurvey_fsw$Email == "Yes", na.rm=TRUE)` 
researchers from the faculty of Social and Behavioural Sciences (FSBS) left their 
email address to be contacted, although not all researchers indicated to be 
willing to meet with us when contacted. So far, `r dim(interviews_fsw)[1]` 
online meetings have been conducted, most of them with a privacy officer present. 
Below, the division over positions can be seen for all interviewees from the 
faculty of Social and Behavioural Sciences (FSBS).

```{r positionsmeetings-fsw}
positionsmeetings(interviews_fsw,
                  title = "Positions one-on-one meetings FSBS",
                  caption = "Positions of researchers of the Faculty of Social and Behavioural Sciences in one-on-one meetings")
```

### Humanities

From the survey respondents, `r sum(dppsurvey_hum$Email == "Yes", na.rm=TRUE)` 
researchers from the Humanities faculty left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews_hum)[1]` online meetings have been 
conducted, all of them with a privacy officer present. Below, the division over 
positions can be seen for all interviewees from the Humanities faculty.

```{r positionsmeetings-hum}
positionsmeetings(interviews_hum,
                  title = "Positions one-on-one meetings Humanities",
                  caption = "Positions of Humanities researchers in one-on-one meetings")
```

### Veterinary Medicine

From the survey respondents, `r sum(dppsurvey_vet$Email == "Yes", na.rm=TRUE)` 
researchers from the faculty of Veterinary Medicine left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews_vet)[1]` online meetings have been 
conducted, most of them with a privacy officer present. Below, the division over 
positions can be seen for all interviewees from the faculty of Veterinary Medicine.

```{r positionsmeetings-vet}
positionsmeetings(interviews_vet,
                  title = "Positions one-on-one meetings Veterinary Medicine",
                  caption = "Positions of researchers from the Faculty of Veterinary Medicine in one-on-one meetings")
```

### LEG

From the survey respondents, `r sum(dppsurvey_leg$Email == "Yes", na.rm=TRUE)` 
researchers from the faculty of Law, Economics, and Governance (LEG) left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews_leg)[1]` online meetings have been 
conducted, most of them with a privacy officer present.  Below, the division over 
positions can be seen for all interviewees from the faculty of Law, Economics, and Governance (LEG).

```{r positionsmeetings-leg}
positionsmeetings(interviews_leg,
                  title = "Positions one-on-one meetings LEG",
                  caption = "Positions of researchers from the Faculty of Law, Economics, and Governance in one-on-one meetings")
```

### Geo

From the survey respondents, `r sum(dppsurvey_geo$Email == "Yes", na.rm=TRUE)` 
researchers from the Geosciences faculty left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews_geo)[1]` online meetings have been 
conducted. Below, the division over 
positions can be seen for all interviewees from the Geosciences faculty.

```{r positionsmeetings-geo}
positionsmeetings(interviews_geo,
                  title = "Positions one-on-one meetings Geosciences",
                  caption = "Positions of Geosciences researchers in one-on-one meetings")
```

### {-}

## Analysis {#analysis}
From the raw Qualtrics output, we first cleaned and split the data into different
data files (cleaned and closed survey responses, open text responses, email
addresses, see the 
<a href="https://github.com/UtrechtUniversity/dataprivacysurvey/blob/main/src/pseudonymise-data.R" 
target="_blank">pseudonymise-data.R</a> script for details). Both the open text 
responses and the notes made during the one-on-one meetings were separately and 
manually scored to enable the extraction of action points. 

Below we report on the descriptive statistics or summaries from most of the 
survey questions. As we did not formulate hypotheses, no statistical analyses 
were performed. 

## Data and material availability {#data-and-material-availability}
All survey-related documentation can be found in the 
<a href="https://github.com/UtrechtUniversity/dataprivacysurvey" target = "_blank">dedicated 
survey repository</a>. It contains:

- The 
<a href = "https://utrechtuniversity.github.io/dataprivacysurvey/assets/survey-questions-qualtrics.pdf"
target = "_blank">full survey</a>
- The survey's <a href = "https://utrechtuniversity.github.io/dataprivacysurvey/assets/survey-privacy-statement.pdf"target = "_blank">privacy statement</a>
- The <a href="https://github.com/UtrechtUniversity/dataprivacysurvey/tree/main/src" target = "_blank">
scripts used to clean the data and write this report</a>

As the dataset contains personal information (demographic information, open text 
responses, email addresses, etc.), and no consent was obtained to share those 
details, we are unable to share them in this repository.

# Results {#results}
## Types of research data {#types-of-research-data .tabset}
### UU-wide

To investigate the types of research that was represented in the sample, we 
asked which types of data, and specifically which types of *personal* data the 
respondents worked with. Most respondents indicated to use tabular, textual, 
code and audio data as their primary research data format. 
In terms of *personal* data types, contact information, demographic information 
and direct identifiers were most common among the respondents. This 
can be either because these are the types of personal data that are indeed most 
common, but possibly also because researchers mostly recognise these types of 
data as being personal data.

```{r datatypesuu}
datatypes(dppsurvey)
```
<br>
    
When comparing faculties, it appears that most personal data is processed in the
faculties of Social and Behavioural Sciences (FSBS), although the faculties of 
Science and Veterinary Medicine also seem to process quite some personal data.
In the tabs, we look further into the types of (personal) data processed within
each faculty.

```{r personaldataplot}
datatypegraphs <- dppsurvey %>% 
  pivot_longer(cols = grep("^Personaldata_type_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Personal_Datatype") %>%
  select(-name) %>%
  pivot_longer(cols = grep("^Faculty_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  mutate(Percentage = round(Count/dim(dppsurvey)[1]*100,1)) %>%
  arrange(Percentage) %>% # Order by percentage
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Percentage)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Faculty, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", 
       title = "Personal data types across Faculties (%)",
       caption = "Percentage of the total sample size of personal data used in every UU faculty") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = paste0(Percentage, "%")),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

datatypegraphs
```

### Science

As with the university-wide data, researchers from the Science faculty indicated
that they also most often used tabular data, textual data and code/theoretical 
models in their research. The same goes for the types of *personal* data: the
same top-3 types are used here as in the entire university (demographic 
information, contact information, direct identifiers). 

```{r datatypesscience}
datatypes(dppsurvey_science)
```
<br>
  
As a large part of the researchers from the Science faculty indicated not to use
*any* personal data whatsoever in their research, it may be useful to 
investigate in which departments these researchers work. In the figure below,
the division of personal data types per Science department is plotted. As can be
expected, the departments of Biology, Physics and Mathematics do not seem to 
process much personal data, whereas in the Information and Computing Sciences
and Pharmaceutical Sciences departments, the most personal data is processed. This
makes sense considering the types of research performed in these departments.

```{r datatypesdeptsscience}
datatypesdepartments(dppsurvey_science, "Science",
                     caption = "Personal data types used in each department of the Science faculty")
```

### FSBS

As with the university-wide data, researchers from the faculty of Social and
Behavioural Sciences (FSBS) indicated that they also most often used tabular data, 
textual data and code/theoretical models in their research. The same goes for 
the types of *personal* data: demographic and contact information are most often
used. Interestingly, sensitive demographic information is used much more relative
to  the UU-wide responses.

```{r datatypesfsw}
datatypes(dppsurvey_fsw)
```
<br>
  
When comparing the types of personal data across departments of the faculty, the
departments of Sociology and Cultural Anthropology seem to process the least
amount of personal data - although these departments may also simply be 
underrepresented in the survey sample. Most other departments process quite some
personal data, as can be expected from a faculty of which the research focuses
almost exclusively on humans. 

```{r datatypesdeptsfsw}
datatypesdepartments(dppsurvey_fsw, "FSW",
                     caption = "Personal data types used in each department of the Faculty of Social and Behavioural Sciences")
```

### Humanities
As with the university-wide data, researchers from the faculty of Humanities 
indicated that they also most frequently used tabular data, textual data and 
audio data in their research. The same goes for the types of *personal* data: the
same top-3 types are present here as in the entire university (demographic 
information, contact information, direct identifiers). 

```{r datatypeshum}
datatypes(dppsurvey_hum)
```
<br>
  
If we look at the different departments within the faculty, it is clear that
mostly the departments of Languages, Literature and Communication, and of Media
and Culture Studies process personal data in their research. Especially 
researchers at the department of Languages, Literature and Communication seem to
process a lot of personal data, which makes sense considering the type of research
performed there.

```{r datatypesdeptshum}
datatypesdepartments(dppsurvey_hum, "Humanities",
                     caption = "Personal data types used in each department of the Faculty of Humanities")
```

### Veterinary Medicine

When comparing the types of data that researchers at the Faculty of Veterinary
Medicine work with, it is clear that researchers at this faculty work with 
slightly different types of data, i.e. biological and physiological data are 
more common among these researchers. Despite this, the most frequent types
of personal data do correspond well to  the university-wide numbers: 
demographic information, direct identifiers and contact information are also 
the most common types of personal data researchers at this faculty seem to 
deal with.

```{r datatypesvet}
datatypes(dppsurvey_vet)
```
<br>
  
When looking at each department separately, it is clear that researchers in the
department of Population Health Sciences process the large majority of the 
personal data within the faculty. Of course, it may also be possible that this 
department was simply more represented in the survey than the other departments.

```{r datatypesdeptsvet}
datatypesdepartments(dppsurvey_vet, "DGK",
                     caption = "Personal data types used in each department of the Faculty of Veterinary Medicine")
```

### LEG
As with the university-wide data, researchers from the faculty of Law, 
Economics and Governance (LEG) indicated that they also most frequently used 
tabular data, textual data and audio data in their research. The same goes for 
the types of *personal* data: the same top-3 types are present here as in the 
entire university (demographic information, contact information, 
direct identifiers). 

```{r datatypesleg}
datatypes(dppsurvey_leg)
```
<br>
  
When looking at the different departments within the faculty, researchers in 
the department of Governance seem to process personal data the most often, 
followed by the department of Law. Of course, it is possible that the other 
departments process more personal data than the respondents indicated. 

```{r datatypesdeptsleg}
datatypesdepartments(dppsurvey_leg, "REBO",
                     caption = "Personal data types used in each department of the Faculty of Law, Economics and Governance")
```

### Geo

As with the university-wide data, researchers from the faculty of Geosciences
indicated that they also most frequently used textual and tabular data. However,
in contrast to the university-wide data, they are followed by geographical data
and code/theoretical models. This is to be expected considering the faculty. In 
terms of *personal* data, however, the same top-3 types are present here as in the 
entire university (demographic information, contact information, 
direct identifiers). 

```{r datatypesgeo}
datatypes(dppsurvey_geo)
```
<br>
  
When comparing departments, it becomes immediately clear that most personal
data seems to be processed in the department of Sustainable Development, and
some also in the Human Geography and Spatial Planning department. This makes 
sense, as the research performed at the third department (Earth sciences) usually
does not focus on human behaviour and thus does not involve much personal data, 
if any. 

```{r datatypesdeptsgeo}
datatypesdepartments(dppsurvey_geo, "Geo",
                     caption = "Personal data types used in each department of the Faculty of Geosciences")
```

## Current practices {#current-practices}

The first part of the survey addressed the researchers' current practices in 
handling personal data in their research.

### Organisational and technical measures {#organisational-technical-measures .tabset}
#### UU-wide
With respect to organisational and technical measures used to handle personal 
data, most respondents indicated that they pseudonymise/anonymise their data. 
As this question is self-reported, we cannot assess whether the researchers' data 
was actually sufficiently pseudonymised/anonymised in line with the GDPR. 
Secondly, many researchers seemed to implement access control and encryption, 
and complete a Data Management Plan (DMP) during their project(s). This is to
be expected, as most funders nowadays require a DMP and many DMP templates 
explicitly address topics like pseudonymisation, access control and encryption. 
On the other hand, GDPR-specific assessments such as Data Protection Impact 
Assessments (DPIAs) or privacy reviews were least used - as it stands, these 
assessments are still only carried out on a case-by-case basis.
 
```{r measuresused}
measures_plot <- 
  dppsurvey %>% 
  pivot_longer(cols = grep("Faculty_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  pivot_longer(cols = grep("^Orgmeasures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Measures") %>%
  select(-name) %>%
  group_by(Measures, Faculty) %>%
  summarise(Count = length(Measures)) %>%
  arrange(Count) %>% # Order by count
  #mutate(Measures = factor(Measures, levels=Measures)) %>% # Update factor levels
  ggplot(aes(reorder(Measures, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Faculty)) +
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  labs(x = "", title = "Protective and planning measures",
       caption = "Organisational and technical measures used in handling personal data") +
  coord_flip() +
  theme_classic() +
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  geom_label(aes(label = after_stat(y), group = Measures),
             stat = 'summary',
             fun = sum,
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             label.size = NA, #1,
             nudge_y = 2,
             label.padding = unit(0.35, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  geom_text(aes(label = Count), # counts per measure per faculty
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))

measures_plot
```


#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Storage media used {#storage-media-used .tabset}
#### UU-wide
When working with personal data, it is important to choose a sufficiently 
secure storage medium. Luckily, most respondents indicated to rely on storage 
solutions that are provided or recommended by UU - and in most cases are indeed
safe for storing personal data. Nonetheless, some respondents indicated to use 
non-UU solutions, including cloud solutions that UU advises against using. 
Several respondents also indicated that they use other storage solutions, such 
as those of external institutions (e.g., University Medical Center, Trimbos 
Institute, Central Bureau of Statistics) and repositories (e.g., DANS EASY, CLARIAH).

```{r storageplot}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Consent forms: usage and content {#consent-forms .tabset}
#### UU-wide
Consent is currently the legal basis most often used in researcher performed at 
UU. Luckily, only a small number of respondents indicated that consent forms 
are not applicable to their research or they don't know what they are. The consent
forms most often used seem to be approved by privacy officers or ethical 
committees. Some respondents however indicated to independently create and use 
consent forms. A handful of respondents indicated to use templates from external 
institutions, or no templates at all (e.g., because they use verbal consent or 
data from another party).

The mere fact that researchers make use of a consent form and information letter 
of course does not mean these forms are of sufficient quality. Therefore, we 
asked respondents what they typically include in their information to 
participants. Most respondents indicated to provide at least the basic 
information on the purposes for data collection and which data will be be used. 
Information on the data subjects' rights and safeguards are often included as 
well. 

Information and consent on data sharing and reuse is apparently not included as 
often. This is a point of concern, because not obtaining consent for sharing 
personal data for reuse may mean that data collected for a specific research
project may not be reused again, reducing the scientific value of the data.

Finally, contact details of the Data Protection Officer (DPO) are not always 
included either, whereas these always have to be provided when working with
personal data (<a href = "https://gdpr-info.eu/art-13-gdpr/" target = "_blank">art. 13</a>)

```{r consentforms}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Data Protection Impact Assessment (DPIA) {#dpia .tabset}
#### UU-wide
A Data Protection Impact Assessment (DPIA) is a legal instrument to assess the
risks involved for data subjects, and helps determine the necessary safeguards to
reduce those risks to an acceptable level. Despite it being an important legal
instrument, most respondents indicated to never have carried one out, or heard of it,
for that matter. A minority of the sample had heard of it, or has completed one. 
Currrently, the desired scenario is that researchers get help from a privacy 
officer when performing a DPIA, and the results seem to indicate that this is
indeed the case in the majority of cases.

```{r dpia}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Data sharing practices {#data-sharing .tabset}
#### UU-wide
To investigate how often data are being shared and under which circumstances, we
asked respondents if and with which parties they typically share their research
data, and which measures they usually take to do so securely. While there were 
some respondents who indicated not to share their data at all, most respondents
seem to only share research data within the organisation, and otherwise within the
European Economic Area (EEA). This is relatively good news, as such transfers 
usually require little, if any, additional safeguards. Responses in the "Other" 
category, however, suggest that a lot of data actually are shared, for example 
with co-authors at another institution, students, or in "pseudonymised form".

Concerning the measures used before sharing data, most repondents indicated to
pseudonymise their data, though we cannot assess the quality of such 
pseudonymisation. Researchers also seemed to use approved tools, agreements, 
and providing access without data transfer. Only a few respondents indicated that 
they involved a data expert while transferring data, and the use of Standard 
Contractual Clauses appears limited.

```{r datasharing}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Data publishing {#data-publishing .tabset}
#### UU-wide

```{r datapublishing}
```

Open science and privacy are often seen as conflicting, as sharing personal data 
cannot be done just like that, but requires at the very least a valid legal 
basis and additional safeguards. Therefore, in our experience, to date not many 
datasets that contain personal data are shared for reuse purposes, and the survey
respondents seemed to confirm this experience, as the majority of the respondents
indicated not to publish their data, or only in anonymised form (again, we cannot
be certain whether the data were indeed entirely anonymised).

The primary reason for not publishing data appears to be that researchers were 
still working on the data, they don't want to/need to publish their data, or 
they cannot anonymise the data. Other reasons given by the respondents included 
that publishing data is too much effort, publication is undesirable, or just not 
considered.

```{r publicationformat}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

## Existing support channels {#existing-support-channels}
The second part of the survey concerned the visibility and use of existing 
support channels. 

### Faculty privacy officer {#faculty-privacy-officer}
When asked whether respondents know who their faculty
privacy officer is, a little over half of the respondents indicated that they do 
not (Yes: `r sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%), No: 
`r sum(dppsurvey$Know_PO=="No", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="No", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%)).

When comparing faculties (see below), it is striking that the faculties where the
most personal data seems to be processed, the majority of respondents is not 
aware of their faculty privacy officer. This suggests either that researchers 
have simply never required help from their privacy officer, or that the faculty 
privacy officers could increase their visibility within their faculties.

```{r knowpoplot}
knowpoplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Faculty_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Faculty") %>%
  select(-name) %>%
  group_by(Faculty, Know_PO) %>%
  summarise(Count = length(Know_PO)) %>%
  arrange(Count) %>% # Order by count
  filter(!is.na(Know_PO)) %>%
  mutate(Know_PO = factor(Know_PO, levels=Know_PO)) %>% # Update factor levels
  ggplot(aes(x = Know_PO, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Faculty, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", 
       title = "Awareness of the privacy officer",
       caption = "Awareness of the privacy officer across UU faculties") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2,
             label.size = NA, #1,
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

knowpoplot

```

### Looking for help {#looking-for-help .tabset}
#### UU-wide
When asked whether respondents have ever looked for support in the form of 
information, tools, or in-person support, an overwheling majority indicated that 
they have, as can be seen in the graph below. Most respondents that looked for support
indicated, however, that they haven't always found the support they were looking for.
Together with the results from the previous question, this suggests that the 
visibility of the current support channels could be improved.

```{r existingsupport}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

### Channels used to find support {#support-channels .tabset}
#### UU-wide
The graph below indicates which channels respondents used most 
when looking for information about handling personal data. Note that this was 
a multiple choice question, and so the numbers do no add up to the total number
of survey respondents. As shown below, all sources mentioned were to some extent
consulted by the respondents. However, there is large differentiation between 
these sources. The university website and intranet are the most visited online 
resources for information about handling personal data. Notably, colleagues 
appear to play an important role as well in informing researchers
about how to handle personal data. This suggests that the better informed 
researchers are, the more positive the effect is on their colleagues as well. 
Moreover, this also suggests that in-person support may be a more effective way 
of increasing awareness of privacy-related practices than more "distant" 
information sources. 

```{r sourcesused}
```

#### Science
Here are results per faculty

#### FBSB
dasd

#### Humanities
asasd

#### Veterinary Medicine
adasds

#### LEG
asasd

#### Geo
asd

### {-}

## Challenges and needs (survey) {#challenges-needs-survey .tabset}

### UU-wide
```{r obstacle}
```

As can be seen above, most researchers experience privacy to be an obstacle for 
open science and research data management in some way. It is therefore important 
to aim for support in this area. What this support should look like, however, 
differs a bit. As can be seen below, accessible information and visible support
channels seem to be the most wanted improvements in the current support, closely
followed by UU-wide policy on the topic, and privacy-related walk-in hours. 

```{r bettersupport}
```

### Science
Here are results per faculty

### FBSB
dasd

### Humanities
asasd

### Veterinary Medicine
adasds

### LEG
asasd

### Geo
asd

## {-}

## Challenges and needs (open questions, meetings) {#challenges-needs-open .tabset}

As mentioned in the [Methods section](#analysis), the responses on the following 
open questions, and the notes taken during the one-on-one meetings with 
researchers were coded to allow for easier analysis. From the survey, the 
following questions were coded:

- "Which challenges concerning the handling of personal data do you run into 
most often?"
- "What specific **information** or **tools** about handling personal data are 
you missing from existing sources?"
- "What can we do better to support you in handling personal data in research?" 
(responses to the "Other" option)

The codes that were assigned in both the survey and the meeting notes can be 
seen in the word cloud below: the larger the font, the more often the code 
was assigned. The meaning of each code is explained in the file 
<a href = "../data/processed/codes-freq_survey_meetings.csv" target = "_blank">"codes-freq_survey_meetings.csv"</a> 
and below. Please note that there is overlap in respondents between the open 
questions in the survey and the meeting notes, and therefore some codes may have 
been applied twice for the same researcher. 

```{r count-codes}
```

```{r count-tools}
```

### UU-wide
```{r wordcloud}
```

### Science
Here are results per faculty

### FBSB
dasd

### Humanities
asasd

### Veterinary Medicine
adasds

### LEG
asasd

### Geo
asd

## {-}
**Below we go into the most frequently mentioned challenges/needs expressed
by researchers in the survey (open questions) and one-one-one meetings:**

### Visibility and findability {#visibility-and-findability}
`r totalcodes$Meaning[totalcodes$Code == "visibility"]` (mentioned 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "visibility"]` times). 
Some respondents indicated not to know where to go for help with privacy-related
matters, others mentioned that there should be *more* support personnel available
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "availability"]`). 
One researcher said to prefer having one place to go to for all
data-related questions. As a matter of fact, many faculties have such a 
"one-stop-shop" in place, but responses such as these indicate that their 
visibility may be improved.

Closely related to visibility is findability. Many researchers point out that 
available information is difficult to find, or it is confusing which source 
should be followed. As an example, one researcher from the faculty of Humanities
was determined that they found some useful information at the RDM Support website.
However, upon closer inspection, it appeared that this information arose from
another data management-related website with the UU logo on it. 

<blockquote>"UU biedt ontzettend veel aan, maar je moet veel websites bezoeken 
om alles te vinden [...] Een duidelijk overzicht voor welke informatie je
waar moet zijn zou fijn zijn."</blockquote>

### Specific tools {#specific-tools}
Some (`r totalcodes$Times_mentioned_total[totalcodes$Code == "tools"]`) 
researchers expressed a need for a concrete (improvement of a) tool, such 
as for:

- `r totaltools$Explanation[totaltools$Name_tool == "transcription"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "transcription"]` 
times)
- `r totaltools$Explanation[totaltools$Name_tool == "computing-critical"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "computing-critical"]` 
times)
- `r totaltools$Explanation[totaltools$Name_tool == "pseudonymising-video"]` (mentioned 
`r totaltools$Times_mentioned_total[totaltools$Name_tool == "pseudonymising-video"]` 
times)

### Hands-on support {#hands-on-support}
Many (`r totalcodes$Times_mentioned_total[totalcodes$Code == "hands-on"]`) 
researchers indicated that support staff could sometimes provide support 
in a more hands-on fashion, rather than abstract advice and telling researchers 
how *not* to do things. Some researchers added that privacy professionals sometimes
have a tendency to cling to the letter of the law, leading to significant
delays in their project, instead of looking at how to concretely solve existing 
issues in practice (mentioned `r totalcodes$Times_mentioned_total[totalcodes$Code == "strict"]` 
times): 

<blockquote>"Soms zijn we door deze regels heiliger dan de paus."</blockquote>
<blockquote>"Actual getting-your-hands-dirty support: not the kind that tells 
you what to do, but also the kind that helps you by doing."</blockquote>

### Less bureaucracy {#less-bureaucracy}
Processes were often experienced as time-inefficient, and sometimes longer 
and more bureaucratic than necessary (mentioned by 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "quicker"]` researchers).
For example, the DPIA process was mentioned explicitly 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "dpia-process"]` times),
as well as having to fill out too many forms with overlapping content (e.g.,
Privacy Scan, Data Management Plan, DPIA). Some 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "divide-responsibilities"]`) 
researchers argued that (part of) this burden should be relieved or 
carried by support staff:

<blockquote>"Minder acties die gericht zijn op inhoudelijk trainen van WP en 
meer uit handen nemen van deze groep."</blockquote>
<blockquote>"Sharing data costs a lot of time and is inefficient when you do 
not do it often"</blockquote>

### Unclear processes and guidelines {#unclear-processes-and-guidelines}
Many (`r totalcodes$Times_mentioned_total[totalcodes$Code == "unclear-process"]`) 
researchers complained that it was unclear what was expected of them when 
they processed personal data in their research, or that they would like to have more,
or more practical guidelines (`r totalcodes$Times_mentioned_total[totalcodes$Code == "guidelines"]`)
on this topic, for example on:

- What steps do researchers need to take? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "assessment-steps"]`)
- Who should researcher ask for help in which situation?
- `r totalcodes$Meaning[totalcodes$Code=="responsibility"]` (`r totalcodes$Times_mentioned_total[totalcodes$Code == "responsibility"]`)
- `r totalcodes$Meaning[totalcodes$Code=="changing-rules"]` (`r totalcodes$Times_mentioned_total[totalcodes$Code == "changing-rules"]`)
- Different processes for different types of research (e.g., student projects vs.
large longitudinal projects, `r totalcodes$Times_mentioned_total[totalcodes$Code == "different-processes"]`)

### Information and education {#information-and-education}
A large part of the respondents expressed a need for more (clear) information 
and education with respect to handling personal data in research. In general, 
researchers indicated that the information offered to them should be 
more clear (`r totalcodes$Times_mentioned_total[totalcodes$Code == "unclear-info"]`), 
simpler `r totalcodes$Times_mentioned_total[totalcodes$Code == "info-simple"]`, 
consistent across resources (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-consistent"]`), 
possibly in the form of templates (`r totalcodes$Times_mentioned_total[totalcodes$Code == "templates"]`). 
Luckily, some researchers indicated to already be happy with existing materials (`r totalcodes$Times_mentioned_total[totalcodes$Code == "positive-existing-materials"]`).

#### Information for specific research or data types {#specific-info}
Many researchers indicated to have a need for more tailored information for specific
types of data or research 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-specific"]`), 
for example for etnographic data 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-etnography"]`),
historical data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-history"]`), 
or video data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "data-video"]`).

<blockquote>"De informatie is gewoon veel te generiek, er zouden templates 
moeten zijn per type onderzoek."</blockquote>
<blockquote>"Veel templates en uitlegmodellen spreken over data, data packages 
en metadata, maar die woorden zijn niet ingebed in historisch onderzoek. Er 
ontstaat al snel verwarring over wat historici nu precies moeten met 
archiefmateriaal in het licht van privacy."</blockquote>

#### Frequently asked questions {#faq}
A selection of researchers used the space in the open questions and/or the meetings 
to ask knowledge-related questions. These can be used in future support efforts,
such as the 
<a href="https://utrechtuniversity.github.io/dataprivacyhandbook" target="_blank">Data 
Privacy Handbook</a>, future courses, or other future information campaigns to 
better address researchers' needs. Below are examples of the most commonly 
asked questions:

- When are data still personal (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-personal-data"]`), 
how to anonymise data (`r totalcodes$Times_mentioned_total[totalcodes$Code == "anonymise-how"]`),
and when are data anonymised sufficiently (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-anonymous"]`)?
- What are the privacy-related requirements for students? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "students"]`)
- How to store different types of personal data? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "datastorage"]`)
- Data sharing: what data can be shared and with whom? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "datasharing"]`)
- How to find a balance between informing data subjects too little vs. providing
too much privacy-related information that will scare them off or hurt their 
trust in research? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "info-informing-toomuch"]`)
- How to balance open science and privacy? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "open-vs-privacy"]`).
Related: when is reusing data fir different purposes allowed? (e.g., education 
data, data collected by students, `r totalcodes$Times_mentioned_total[totalcodes$Code == "info-reuse"][1]`)
- How to collaborate with multiple institutions? (`r totalcodes$Times_mentioned_total[totalcodes$Code == "collaboration"]`)

#### Education and onboarding {#education-and-onboarding}
The need for more educational resources was also recognised by a selection of the
respondents. Concretely, researchers could be educated more in the following ways:

- Privacy and/or research data management as part of the master or PhD 
curriculum (`r totalcodes$Times_mentioned_total[totalcodes$Code == "education"]`)
- Privacy as part of the onboarding procedure of new employees (`r totalcodes$Times_mentioned_total[totalcodes$Code == "onboarding"]`)
- Mandatory privacy training for supervisors, principal investigators, professors, 
and/or teachers (`r totalcodes$Times_mentioned_total[totalcodes$Code == "education-supervisors"]`)
- A course on how to handle personal data in research (`r totalcodes$Times_mentioned_total[totalcodes$Code == "course"]`)

<blockquote>"There is no one who tells at the start of your PhD how you should handle your data. [...] I think new PhD students should get a basic course on data management and privacy."</blockquote>

### ... or simply no issues {#no-issues}
Notably, there were also researchers who indicated **not** to have run into issues 
(yet), or to have received sufficient and useful help (`r totalcodes$Times_mentioned_total[totalcodes$Code == "positive-existing-support"]`). For example:

<blockquote>"The data manager and privacy officer of the faculty of humanities help a lot. This support is essential!"</blockquote>
<blockquote>"Tot nu toe heb ik niet veel problemen gehad. De institutional review board van onze afdeling kijkt altijd kritisch naar de onderzoeksvoorstellen, ook met name op omgaan met persoonsgegevens."</blockquote>

# Technical information {#technical-information}
This document was last created on: `r as.character(Sys.Date())`. It was 
created in the following environment:

```{r sessioninfo}
sessionInfo()
```
