---
title: "Data Privacy Survey"
subtitle: "Results for the Science faculty"
author:
- "Dorien Huijser"
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
abstract: |
  In the second quarter of 2022, Utrecht University (UU) Research Data Management 
  Support (RDM Support) sent out a survey among all scientific personnel at UU: 
  the Data Privacy Survey. The aim of this survey was to investigate 1) How UU 
  researchers currently deal with personal data in research, 2) What challenges 
  they run into when handling personal data in research, and 3) How support at
  UU can improve their services concerning personal data in research. The current
  report summarises the results from respondents from the Science faculty. 
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
    df_print: paged
    css: style.css
    includes:
      in_header: "../includes/header.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

```{r sourcescript, include=FALSE, cache=FALSE}
knitr::read_chunk('plot-data.R')
```

# Introduction {#introduction}
The Data Privacy Survey was conducted in 2022 at Utrecht University (UU) Research
Data Management Support (RDM Support), as part of a larger project, the
<a href="https://utrechtuniversity.github.io/dataprivacyproject" target = "_blank">Data Privacy Project</a>. 
It aimed to find out:

1. To what extent are UU researchers aware of privacy legislation and practices?
2. What data privacy issues do UU researchers typically run into?
3. What support do UU researchers need to handle personal data in their research?

These questions were investigated through an online survey and subsequent
one-on-one meetings with researchers, who indicated to be willing to speak with us
about the topic of handling personal data in research. The full report of this
survey can be read in the 
<a href = "data-privacy-survey-report-v0.2.Rmd">university-wide report</a>.

As every UU faculty is organised somewhat differently, privacy-related 
experiences of UU researchers may also depend heavily on the faculty where they 
perform their research. Therefore, here we created an additional report for
the Science faculty, the largest faculty at the university. 

# Methods {#methods}
The full methods are described in the UU-wide report. Below is a summary of the
details specific to the Science faculty.

```{r, dependencies}
```

```{r readdata}
```

```{r selectscience}
# Select only Science respondents
dppsurvey <- dppsurvey %>% 
  filter(!is.na(Faculty_5))
```

## Online survey respondents {#online-survey-respondents}
There were `r dim(dppsurvey)[1]` respondents from the Science faculty in the 
online survey. It took them a median of `r round(median(dppsurvey$Duration/60), 1)` minutes 
to complete it (ranging from `r round(min(dppsurvey$Duration)/60, 1)` minutes to 
`r round(max(dppsurvey$Duration)/3600/24,1)` days). In the graph below,
the representation of each department within the Science faculty is visualised. 

```{r uucolors}
```

```{r departmentscience}
# Plot of positions per faculty
dppsurvey %>%
  pivot_longer(cols = grep("Dept_Science_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  pivot_longer(cols = grep("Position_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Position") %>%
  select(-name) %>%
  group_by(Department, Position) %>%
  summarise(Count = length(Department)) %>%
  ggplot(aes(reorder(Department, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Position)) +
  coord_flip() +
  theme_classic() + 
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  labs(x = "", 
       title = "Department and Position representation",
       caption = "Survey respondents per department and position in the Science faculty") +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  #scale_x_continuous(expand = expansion(add = 3)) + 
  geom_label(aes(label = after_stat(y), group = Department), # totals per dept
             stat = 'summary', 
             fun = sum, 
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             nudge_y = 2,
             label.size = NA, #1,
             label.padding = unit(0.35, "lines"),
             #label.r = unit(0.25, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  geom_text(aes(label = Count), # counts per position per department
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))
```

Most respondents in the online survey seemed to come from the Information and
Computing Sciences (ICS) and Pharmaceutical Sciences departments, and were 
relatively early career researchers (PhD/junior/postdoc researcher, 
assistant/associate professor). This makes sense, as more personal data is 
likely processed in the fields of ICS and pharmaceuticals, compared to, for 
example, mathematics.

## One-on-one meetings {#one-on-one-meetings}
```{r read-opentext}
```

```{r coded-science}
# Filter interview and open text responses codes on Science faculty
interviews <- interviews %>%
    filter(grepl("Science(?!s)", Faculty, perl = TRUE))

opentext <- opentext %>%
      filter(grepl("Science(?!s)", Faculty, perl = TRUE))
```

From the survey respondents, `r sum(dppsurvey$Email == "Yes", na.rm=TRUE)` 
researchers from the Science faculty left their email address to be contacted,
although not all researchers indicated to be willing to meet with us when 
contacted. So far, `r dim(interviews)[1]` online meetings have been 
conducted, each taking about 30 minutes, and each with the project coordinators. 

Below, the division over positions can be seen for all interviewees from the 
Science faculty.

```{r positionsmeetings}
```

# Results {#results}
## Types of research data {#types-of-research-data}
As with the university-wide data, researchers from the Science faculty indicated
that they also most often used tabular data, textual data and code/theoretical 
models in their research. The same goes for the types of *personal* data: the
same top-3 types are used here as in the entire university (demographic 
information, contact information, direct identifiers). 

```{r datatypes}
```

As a large part of the researchers from the Science faculty indicated not to use
*any* personal data whatsoever in their research, it may be useful to 
investigate in which departments these researchers work. In the figure below,
the division of personal data types per Science department is plotted. As can be
expected, the departments of Biology, Physics and Mathematics do not seem to 
process much personal data, whereas in the Information and Computing Sciences
and Pharmaceutical Sciences departments, the most personal data is processed. This
makes sense considering the types of research performed in these departments.

```{r beta-personaldataplot}
beta_datatypegraphs <- personaldatatypes %>%
  pivot_longer(cols = grep("^Dept_Science_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Personal_Datatype) %>%
  summarise(Count = length(Personal_Datatype)) %>%
  arrange(Count) %>% # Order by count
  mutate(Personal_Datatype = factor(Personal_Datatype, levels=Personal_Datatype)) %>% # Update factor levels
  ggplot(aes(x = Personal_Datatype, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", title = "Personal data types across Science departments") +
  coord_flip() +
  theme_classic() +
  geom_text(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2, #"inward",
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

beta_datatypegraphs
```

## Current practices {#current-practices}

The first part of the survey addressed the researchers' current practices in 
handling personal data in their research.

### Organisational and technical measures {#organisational-technical-measures}
With respect to organisational and technical measures used to handle personal 
data, most respondents in the Science faculty indicated that they 
pseudonymise/anonymise their data, use access control, and fill out a Data 
Management Plan (DMP). These results are comparable to the university-wide results. 

Notably, most responses to this and upcoming questions seem to be filled out
primarily by researchers from the ICS and Pharmaceutical Sciences departments.
It is possible that researchers from the other departments did not answer these
questions, as they do not process a lot of personal data in their research, and
may have simply skipped these questions.
 
```{r measuresused}
dppsurvey %>% 
  pivot_longer(cols = grep("Dept_Science_[0-9]$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  pivot_longer(cols = grep("^Orgmeasures_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Measures") %>%
  select(-name) %>%
  group_by(Measures, Department) %>%
  summarise(Count = length(Measures)) %>%
  arrange(Count) %>% # Order by count
  #mutate(Measures = factor(Measures, levels=Measures)) %>% # Update factor levels
  ggplot(aes(reorder(Measures, 
                     -Count, 
                     sum, 
                     decreasing = T), 
             Count, 
             fill = Department)) +
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  labs(x = "", title = "Protective and planning measures",
       caption = "Organisational and technical measures used in handling personal data") +
  coord_flip() +
  theme_classic() +
  geom_col() +
  scale_fill_manual(values = UU_pallette) +
  scale_y_discrete(expand = expansion(add = 4)) + # make field larger to see label
  geom_label(aes(label = after_stat(y), group = Measures),
             stat = 'summary',
             fun = sum,
             vjust = 0.5, 
             hjust = -0.2, #"inward",
             label.size = NA, #1,
             nudge_y = 2,
             label.padding = unit(0.15, "lines"),
             color = "black",
             fill = "#FFFFFF") + #"#FFCD00") +
  geom_text(aes(label = Count), # counts per measure per dept
            color = "white",
            fill = "#FFFFFF",
            position = position_stack(vjust = 0.5)) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 11),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank()) +
  guides(fill = guide_legend(nrow = 3, byrow = FALSE))
```

### Storage media used {#storage-media-used}
The majority of respondents from the Science faculty indicated to use a UU-provided or
-approved storage medium (both cloud and network solutions) for storing
personal data. This is good news as these are the facilities that are safe to 
use. As with the university-wide results, other facilities were also sometimes 
used, including those not recommended by UU. Respondents who filled out those
non-recommended responses got a warning to take a critical look at their
choice of storage media before moving on to the next survey question. In one
of the one-on-one interviews with a Science researcher, we learned that this
researcher had taken this advice from the warning to heart and had switched to a 
UU-recommended storage facility instead. 

```{r storageplot}
```

### Consent forms: usage and content {#consent-forms}

In the below graph the responses to the questions about consent forms are 
displayed. They make clear that most researchers at the Science faculty use a 
consent form that is approved by a privacy expert or ethical committee, but also
that part of the respondents did not use a consent form at all, or wrote it 
themselves. This is a bit worrying, as consent is at the moment of writing the
legal basis most often used for research performed at UU, and writing one may 
take expertise that researchers do not always have (nor do they need to have it).

When a consent form is used, the information provided to data subjects does 
seem to be relatively complete, as indicated by the division of responses to this 
question: many components of this (multiple responses) question seem to be present.
Only contact details of the Data Protection Officer (DPO) and information and
consent for further processing and sharing the data for reuse is not regularly
included. 

```{r consentforms}
```

### Data Protection Impact Assessment (DPIA) {#dpia}

There does not seem to be much experience with performing DPIAs at the Science
department: most researchers indicated not to have performed any, nor to have
heard from it. Those that did perform a DPIA did not always get help from a 
privacy officer. This could be interpreted as problematic, however, it is also
possible that there was no privacy officer available yet when these researchers
performed their DPIA, or that they got help from someone with similar expertise,
but without the title of "privacy officer". 

```{r dpia}
```

### Data sharing practices {#data-sharing}

It appears that data are not much shared externally by researchers within the 
Science faculty. Those that did indicate to share data, used mostly 
pseudonymisation, approved tools, and agreements to protect the personal data. 

```{r datasharing}
```

### Data publishing {#data-publishing}

```{r datapublishing}
```
Datasets containing personal data also do not seem to be published at all, or
only the metadata (i.e., contextual information about the project, which usually
only contains personal information from the researchers themselves). The main
reasons for this seems to be that researchers simply felt they did not need to
do so, or (thought that they) had no legal basis to do so. 

```{r publicationformat}
```

## Existing support channels {#existing-support-channels}
### Faculty privacy officer {#faculty-privacy-officer}
When asked whether respondents know who their faculty
privacy officer is, more than half of the respondents indicated that they do 
not (Yes: `r sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="Yes", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%), No: 
`r sum(dppsurvey$Know_PO=="No", na.rm=TRUE)` 
(`r round(sum(dppsurvey$Know_PO=="No", na.rm=TRUE)/sum(!is.na(dppsurvey$Know_PO))*100,0)`%)).

When comparing departments (see below), it is striking that even in the 
departments where most personal data are processed in research, a large part of
the respondents do not seem to know the faculty privacy officer. This suggests 
either that researchers have simply never required help from their privacy 
officer, or that the faculty privacy officer could increase their visibility 
within the faculty.

```{r knowpoplot}
knowpoplot <- dppsurvey %>% 
  pivot_longer(cols = grep("^Dept_Science_[0-9]+$", names(dppsurvey), value=TRUE), 
               values_drop_na = TRUE,
               values_to = "Department") %>%
  select(-name) %>%
  group_by(Department, Know_PO) %>%
  summarise(Count = length(Know_PO)) %>%
  arrange(Count) %>% # Order by count
  filter(!is.na(Know_PO)) %>%
  mutate(Know_PO = factor(Know_PO, levels=Know_PO)) %>% # Update factor levels
  ggplot(aes(x = Know_PO, y = Count)) + 
  geom_bar(stat = "identity", fill = "#FFCD00") + 
  facet_wrap(~Department, ncol = 4,
             labeller = label_wrap_gen(width=14)) +
  labs(x = "", 
       title = "Awareness of the privacy officer",
       caption = "Awareness of the privacy officer across Science departments") +
  coord_flip() +
  theme_classic() +
  geom_label(aes(label = Count),
             #vjust = 0.5, 
             hjust = "inward", #-0.2,
             label.size = NA, #1,
             color = "black") +
  theme(axis.title.x = element_text(size = 10),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks = element_blank(),
        #panel.background = element_rect(fill = NA, color = "black"),
        panel.background = element_blank(),
        panel.border=element_blank(), #rect(fill = NA),
        panel.spacing.y = unit(1, "lines"),
        panel.grid.major.y = element_line(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        strip.background = element_rect(color="white",
                                        size=1.5),
        strip.text = element_text(family = "Verdana",
                                  size = 10,
                                  face = "bold"))

knowpoplot
```

### Looking for help {#looking-for-help}
When asked whether respondents have ever looked for support in the form of 
information, tools, or in-person support, most researchers in the Science faculty
indicated that they have, as can be seen in the graph below. Luckily, most 
respondents that looked for support indicated to also have found what they were
looking for. This suggests that the visibility of the current information,
support and tools at the Science faculty is already relatively good when 
compared to the university-wide results.

```{r existingsupport}
```

### Channels used to find support {#support-channels}
The graph below indicates which channels respondents used most 
when looking for information about handling personal data. Strikingly, 
researchers at the Science faculty seemed to rely most on direct colleagues,
rather than available trainings and specific websites. This is in contrast with
the university-wide results, where the UU website and intranet scored the 
highest. This suggests that in-person support may be a more effective way 
of increasing awareness of privacy-related practices than more "distant" 
information sources. 

```{r sourcesused}
```

## Challenges and needs (survey) {#challenges-needs-survey}

```{r obstacle}
```

As can be seen above, most Science researchers experience privacy to be an 
obstacle for open science and research data management in some way. It is 
therefore important to aim for support in this area. According to the respondents
in the Science faculty, this support would be best improved by increasing the
visibility of available support, improving or increasing the information about
handling personal data in research, and drawing up a UU-wide policy on handling
personal data in research. 

```{r bettersupport}
```

## Challenges and needs (open questions, meetings) {#challenges-needs-open}

The codes that were assigned in both the survey and the meeting notes can be 
seen in the word cloud below: the larger the font, the more often the code 
was assigned. Please note that there is overlap in respondents between the open 
questions in the survey and the meeting notes, and therefore some codes may have 
been applied twice for the same researcher. 

```{r count-codes}
```

```{r count-tools}
```

```{r wordcloud}
```

**Below we go into the most frequently mentioned challenges/needs expressed
by researchers at the Science faculty in the survey (open questions) and 
one-one-one meetings:**

### Specific tools {#specific-tools}
The need for specific tools was mentioned by 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "tools"]` Science 
researchers. There was however wide variation in the type of tools requested.
Most tools related to pseudonymising data in some way, for example qualitative 
data or video data. Here's the top 3:

- `r str_to_sentence(totaltools$Explanation[1])` (mentioned 
`r totaltools$Times_mentioned_total[1]` times)
- `r str_to_sentence(totaltools$Explanation[2])` (mentioned 
`r totaltools$Times_mentioned_total[2]` times)
- `r str_to_sentence(totaltools$Explanation[3])` (mentioned 
`r totaltools$Times_mentioned_total[3]` times)

Besides these tools, a selection of researchers also indicated that they would
like to be able to use templates for frequent privacy-related documentation
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "templates"]` times).
One example would be a template DPIA, pre-filled for specific types of research, which
requires only minimal effort to adjust to a specific research project.

### Hands-on support {#hands-on-support}
Similar to the university-wide results, a selection of Science researchers 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "hands-on"]`)
also indicated to have a need for more hands-on support, that is support that helps
researchers more concretely and clings less to the letter of the law (mentioned `r totalcodes$Times_mentioned_total[totalcodes$Code == "strict"]` 
times)in favor of providing practical solutions.

### Visibility and availability {#visibility-availability}
Visibility of the available support was also mentioned regularly by Science 
researchers (visibility mentioned 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "visibility"]` times, 
availability `r totalcodes$Times_mentioned_total[totalcodes$Code == "availability"]`
times). Combined with the fact that Science researchers seemed to ask their direct 
colleagues a lot when looking for support, increasing the awareness of available
data management and privacy-related expertise may need to be prioritised.

### Less bureaucracy {#less-bureaucracy}
Some researchers at the Science faculty have apparently also faced some 
bureaucracy and slow processes in handling personal data in their research. These
researchers mentioned that some processes should be made easier and more efficient
(mentioned by `r totalcodes$Times_mentioned_total[totalcodes$Code == "quicker"]` 
researchers), especially that of performing a DPIA  (mentioned explicitly 
`r totalcodes$Times_mentioned_total[totalcodes$Code == "dpia-process"]` times). 

### ... or simply no issues {#no-issues}
Notably, there were also a few researchers who indicated **not** to have run 
into issues (yet), or to have received sufficient and useful help 
(`r totalcodes$Times_mentioned_total[totalcodes$Code == "no-issues"]`).

# Conclusion {#conclusion}
In this report, we summarised the findings of the Data Privacy Survey, looking
specifically at respondents from the faculty of Science. In short, we found that:

- Most personal data are processed in the departments of Information Computing 
Sciences and Pharmaceutical Sciences. The types of data are similar to that of 
the entire university, namely tabular and textual data and code/theoretical 
models. In terms of personal data, demographic information, direct identifiers 
and contact information were most common.
- The current practices of researchers at the Science faculty resemble those 
explicated in the university-wide results: most researchers are familiar with the
basic concepts and most common practices (e.g., DMPs, informed consent forms, 
anonymisation), but other, also important, practices were less often performed, 
such as the DPIA, data publication, and measures such as storage encryption or 
the processing register.
- As there hasn't been a privacy officer for that long of a time at the Science 
faculty, it is not that surprising that not all researchers are yet familiar 
with them. It follows that most researchers turn to each other or the internet
to look for information and support. Increasing the awareness of available support,
especially in the ICS and Pharmaceutical sciences departments, may be fruitful
in improving the compliance of researcher performed at this faculty. 

Some recommendations are already described in the section above on 
[challenges and needs](##challenges-needs-open). For further recommendations, 
please refer to the 
<a href="data-privacy-survey-report-v0.2.Rmd" target = "_blank">university-wide report</a>.

# Technical information {#technical-information}
This document was last created on: `r as.character(Sys.Date())`. It was 
created in the following environment:

```{r sessioninfo}
```